---
title: "Extract Info From Data Dictionary"
author: "Simran Makwana & Alba Gutierrez"
date: "11/23/2021"
output: html_document
---

Goal: Save lab dictonary containing mapping frome eCRF to BCH concept_cd. 

# Begin with Exploring the entire Lab Values Data Dictionary
```{r setup, include=FALSE}
library(tidyverse)


# for testing
bch_data <- read.delim('../local_ref/music_labs_observation_withUnits.dsv')
```

Load data
```{r}
datadict <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv')
checked_vars <- read.csv('../data/MUSIC Laboratory Values Form_ mapping from eCRF numeric variables and BCH variables - eCRF to BCH laboratory numeric variables.csv', skip = 1) 
```

*Focus on all values possible*
Focus on values which have passed the checkUnits filters.
```{r}
#var_prefixes <- checked_vars %>% 
#  filter(checkUnits == 'OK') %>%
#  pull(Variable...Field.Name) 

#var_prefixes <- gsub('_value', '', var_prefixes)

# focusing on these 4 types of variables first
#our_vars <- c(paste0(var_prefixes, '_obtained'), 
#              paste0(var_prefixes, '_value'),
#              paste0(var_prefixes, '_unit'),
#              paste0(var_prefixes, '_date'))

  
```

```{r}


var_func <- function(variablename){
  
  for (prefix in var_prefixes) {
    if (grepl(prefix, variablename)) {return(paste0(prefix, '_value'))}
  }
  
  return('')
  
}


```



Focus on laboratory values in data dictionary
- combine the "Choices..." column with the "Field.Note" column, as they contain similar info
- merge in the data from our manual check with bch (checked_vars)
```{r}
datadict_labonly <- datadict %>% 
  filter(Form.Name == "laboratory_values") %>%
  mutate(Choices..Calculations..OR.Slider.Labels = paste0(Choices..Calculations..OR.Slider.Labels, Field.Note)) 

value_vars <- datadict_labonly$Variable...Field.Name[grepl('_value$', datadict_labonly$Variable...Field.Name)]
var_prefixes <- gsub('_value', '', value_vars)

datadict_labonly <- datadict_labonly %>%
  mutate(var_match = sapply(Variable...Field.Name, var_func)) %>%
  left_join(checked_vars %>% 
              select(variable, concept_cd, name_char, concept_path, checkUnits), 
            by = c('var_match' = 'variable')) %>%
  select(-Field.Type, -Field.Label, -Form.Name, -Branching.Logic..Show.field.only.if...., 
         -Required.Field., -Custom.Alignment, -Question.Number..surveys.only., 
         -Matrix.Group.Name, -Matrix.Ranking., -Field.Annotation) %>%
  unique()

datadict_labonly
```

### Fill in missing values

0. We have some labs which we received feedback on but could not find. incorporate those here. 
```{r}
datadict_labonly <- datadict_labonly %>%
  add_row(Variable...Field.Name = 'lactate_obtained',
          concept_cd = 'LAB:3775554',
          name_char = 'Lactic Acid, Whole Blood',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\Lactic Acid, Whole Blood\\LAB:3775554\\",
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'lactate_date',
          concept_cd = 'LAB:3775554',
          name_char = 'Lactic Acid, Whole Blood',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\Lactic Acid, Whole Blood\\LAB:3775554\\",
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'lactate_value',
          concept_cd = 'LAB:3775554',
          name_char = 'Lactic Acid, Whole Blood',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\Lactic Acid, Whole Blood\\LAB:3775554\\",
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'lactate_unit',
          concept_cd = 'LAB:3775554',
          name_char = 'Lactic Acid, Whole Blood',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\Lactic Acid, Whole Blood\\LAB:3775554\\",
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'urea_obtained',
          concept_cd = 'LAB:3775158',
          name_char = 'BUN',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\BUN\\LAB:3775158\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'urea_date',
          concept_cd = 'LAB:3775158',
          name_char = 'BUN',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\BUN\\LAB:3775158\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'urea_value',
          concept_cd = 'LAB:3775158',
          name_char = 'BUN',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\BUN\\LAB:3775158\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'urea_unit',
          concept_cd = 'LAB:3775158',
          name_char = 'BUN',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\BUN\\LAB:3775158\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'bicarbonate_obtained',
          concept_cd = 'LAB:3710038',
          name_char = 'CO2',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\CO2\\LAB:3710038\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'bicarbonate_date',
          concept_cd = 'LAB:3710038',
          name_char = 'CO2',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\CO2\\LAB:3710038\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'bicarbonate_value',
          concept_cd = 'LAB:3710038',
          name_char = 'CO2',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\CO2\\LAB:3710038\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'bicarbonate_unit',
          concept_cd = 'LAB:3710038',
          name_char = 'CO2',
          concept_path = "\\i2b2\\Lab View\\Laboratory\\Chemistry\\CO2\\LAB:3710038\\",
          checkUnits = "Expected and current units match") %>%
  add_row(Variable...Field.Name = 'sars_cov2ntpcr_obtained',
          concept_cd = 'LAB:1044704735',
          name_char = 'SARS CoV-2 (COVID-19) PCR, Resp, QuaL',
          concept_path = NA,
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'sars_cov2ntpcr_date',
          concept_cd = 'LAB:1044704735',
          name_char = 'SARS CoV-2 (COVID-19) PCR, Resp, QuaL',
          concept_path = NA,
          checkUnits = 'Expected and current units match') %>%
  add_row(Variable...Field.Name = 'sars_cov2ntpcr_value',
          concept_cd = 'LAB:1044704735',
          name_char = 'SARS CoV-2 (COVID-19) PCR, Resp, QuaL',
          concept_path = NA,
          checkUnits = 'Expected and current units match')  

```


1. Variables which are 'good to go'
  - checkUnits is ok
  - at least one concept_cd per varname
```{r}
finalmap <- datadict_labonly %>% 
  filter(checkUnits == 'Expected and current units match',
         concept_cd != '',
         !is.na(concept_cd)) %>%
  select(-Choices..Calculations..OR.Slider.Labels, -Field.Note, -Text.Validation.Type.OR.Show.Slider.Number, 
         -Identifier., -var_match, -Text.Validation.Min, -Text.Validation.Max, -Section.Header)

print(paste0('There are ', length(unique(finalmap$Variable...Field.Name)), " variables which are 'good to go'"))

```
  
2. Missing concept_cd
  - find one manually
```{r}
missing_cd <- datadict_labonly %>% 
  filter(concept_cd == '' | is.na(concept_cd))

print(paste0('There are ', length(unique(missing_cd$Variable...Field.Name)), " variables which are have missing concept_cd"))

missing_cd

# of these, we KNOW we can remove the following:
remove <- c('lab_values_na', # we cannot know this from auto extraction
            'lab_values_visit', # we calculate this later
            'anti_sars_igm_man', # we cannot know this from auto extraction
            'anti_sars_igg_man', # we cannot know this from auto extraction
            'total_anti_man', # we cannot know this from auto extraction
            'anc', # this will be calculated by redcap
            'alc', # this will be calculated by redcap
            'nl_ratio', # this will be calculated by redcap
            'cytokine_obtained', # there is no concept related to this. we COULD determine based on other lab values
            'other_il1_obtained', # we cannot know this from auto extraction
            'other_il1_date', # we cannot know this from auto extraction
            'other_il1_value', # we cannot know this from auto extraction
            'other_il1_unit', # we cannot know this from auto extraction
            'other_il2_obtained', # we cannot know this from auto extraction
            'other_il2_date', # we cannot know this from auto extraction
            'other_il2_value', # we cannot know this from auto extraction
            'other_il2_unit', # we cannot know this from auto extraction
            'other_il3_obtained', # we cannot know this from auto extraction
            'other_il3_date', # we cannot know this from auto extraction
            'other_il3_value', # we cannot know this from auto extraction
            'other_il3_unit', # we cannot know this from auto extraction
            'other_il4_obtained', # we cannot know this from auto extraction
            'other_il4_date', # we cannot know this from auto extraction
            'other_il4_value', # we cannot know this from auto extraction
            'other_il4_unit', # we cannot know this from auto extraction
            'other_il5_obtained', # we cannot know this from auto extraction
            'other_il5_date', # we cannot know this from auto extraction
            'other_il5_value', # we cannot know this from auto extraction
            'other_il5_unit', # we cannot know this from auto extraction
            'other_labs', # we cannot know this from auto extraction
            'other1_specify', # we cannot know this from auto extraction
            'other1_date', # we cannot know this from auto extraction
            'other1_value', # we cannot know this from auto extraction
            'other1_unit', # we cannot know this from auto extraction 
            'other2_specify', # we cannot know this from auto extraction
            'other2_date', # we cannot know this from auto extraction
            'other2_value',  # we cannot know this from auto extraction
            'other2_unit', # we cannot know this from auto extraction
            'other3_specify', # we cannot know this from auto extraction
            'other3_date',  # we cannot know this from auto extraction
            'other3_value', # we cannot know this from auto extraction
            'other3_unit', # we cannot know this from auto extraction
            'other4_specify', # we cannot know this from auto extraction 
            'other4_date', # we cannot know this from auto extraction
            'other4_value',  # we cannot know this from auto extraction
            'other4_unit', # we cannot know this from auto extraction
            'other5_specify', # we cannot know this from auto extraction
            'other5_date', # we cannot know this from auto extraction
            'other5_value', # we cannot know this from auto extraction
            'other5_unit' # we cannot know this from auto extraction
            ) 

missing_cd <- missing_cd %>% filter(!Variable...Field.Name %in% remove)
missing_cd
print(paste0('There are now ', length(unique(missing_cd$Variable...Field.Name)), " variables which are have missing concept_cd"))

datadict_labonly <- datadict_labonly %>% filter(!Variable...Field.Name %in% remove)

# we manually review the remaining variables and find the following information (which we add to the final mapping table)
# we got this information from mohamad as LabCode in the 59 account
# we want to add this to finalmap once (if) verified

datadict_labonly <- datadict_labonly %>% 
  add_row(Variable...Field.Name = 'anti_sars_igm_obtained',
          concept_cd = 'LAB:1055858187',
          name_char = 'SARS CoV-2 IgM',
          concept_path = NA,
          checkUnits = 'Our patients have values TNP and L, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'anti_sars_igm_date',
          concept_cd = 'LAB:1055858187',
          name_char = 'SARS CoV-2 IgM',
          concept_path = NA,
          checkUnits = 'Our patients have values TNP and L, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'anti_sars_igm_value',
          concept_cd = 'LAB:1055858187',
          name_char = 'SARS CoV-2 IgM',
          concept_path = NA,
          checkUnits = 'Our patients have values TNP and L, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'anti_sars_igg_obtained',
          concept_cd = 'LAB:1055858179',
          name_char = 'SARS CoV-2 IgG',
          concept_path = NA,
          checkUnits = 'Our patients have numeric values with NOT DEFINED IN SOURCE units, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'anti_sars_igg_date',
          concept_cd = 'LAB:1055858179',
          name_char = 'SARS CoV-2 IgG',
          concept_path = NA,
          checkUnits = 'Our patients have numeric values with NOT DEFINED IN SOURCE units, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'anti_sars_igg_value',
          concept_cd = 'LAB:1055858179',
          name_char = 'SARS CoV-2 IgG',
          concept_path = NA,
          checkUnits = 'Our patients have numeric values with NOT DEFINED IN SOURCE units, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'total_anti_obtained',
          concept_cd = 'LAB:1062063907',
          name_char = 'SARS CoV-2 Total Antibody',
          concept_path = NA,
          checkUnits = 'Our patients have Non-reactive and Reactive, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'total_anti_date',
          concept_cd = 'LAB:1062063907',
          name_char = 'SARS CoV-2 Total Antibody',
          concept_path = NA,
          checkUnits = 'Our patients have Non-reactive and Reactive, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate') %>%
  add_row(Variable...Field.Name = 'total_anti_value',
          concept_cd = 'LAB:1062063907',
          name_char = 'SARS CoV-2 Total Antibody',
          concept_path = NA,
          checkUnits = 'Our patients have Non-reactive and Reactive, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate')

 
  

```

3. Incorrect or missing checkUnits
- we add IgE to the lab dictionary, as we account for the difference in units in the lab_value_etl_with_pre_existing_follow_up script
```{r}
test <- datadict_labonly %>%
  group_by(Variable...Field.Name) %>%
  summarise(concept = paste0(concept_cd, collapse = '-'),
            checkUnits = paste0(checkUnits, collapse = '-'))

testgood <- test %>% filter(concept != 'NA' | concept != '', grepl('match', checkUnits))
testbad <- test %>% filter(concept == 'NA' | concept == '' | !grepl('match', checkUnits))

testbad

######
# we can still use the variables where the units do not match, but just need to treat it differently in the redcap output (list unit, etc)
finalmap <- rbind(finalmap, 
            datadict_labonly %>% filter(grepl('ige', Variable...Field.Name)) %>% select(Variable...Field.Name, concept_cd, name_char, concept_path, checkUnits))

```

```{r}
#loinc_bch <- read.csv('../data/BCH_Lab_Loinc_cd_Map_data.csv')
#ecrf_loinc <- read.csv('../data/eCRF_to_LOINC_forReview.csv')
```


4. checkUnits unverified due to not having MISC patient data
```{r}

# If there are no data found for these patients in BCH, but we have a concept_cd, we cannot do anything.
# If BCH expects to find labs for this patient, they should provide us with a list of keywords which might be associated with the lab.
testbad2 <- testbad %>% filter(grepl('no data for MIS-C patients in BCH', checkUnits))



```

There are still some variables we need to remove.
```{r}
# We remove crp_unit_other, wbc_unit_other, platelets_unit_other, platelets_unit_other  because our units match the specified ones.
finalmap <- finalmap %>% filter(!Variable...Field.Name %in% c('crp_unit_other', 'wbc_unit_other', 'platelets_unit_other', 'triglycerid_unit_other') )

```

We now have all the variables which we can map reported value in the EHR to expected value in RedCap.
```{r}
labdict <- finalmap %>% 
  select(Variable...Field.Name, concept_cd) %>%
  rename('variableName' = Variable...Field.Name) %>%
  mutate(formName = 'laboratory_values')

```

Save labdict as `labdictionary.txt` for use in other notebooks.
```{r}
write.table(labdict, '../data/labDictionary_updated.txt', quote = FALSE, row.name = FALSE)

# for manual SQL query:
#write.table(labdict$concept_cd, '../local_ref/concepts_for_query_updated.txt', row.names = FALSE, col.names = FALSE, eol = "', '", quote = FALSE)

```

```{r}
sessionInfo()
```








