---
title: 'Lab Values: Data ETL'
author: "Simran Makwana & Alba Gutierrez"
date: "11/24/2021"
output: html_document
---

*derived from `music_lab_Values.R`*


```{r}
library(tidyverse)
library(lubridate)
#lab_dictionary <- read.delim("../data/labDictionary_updated.txt", header = TRUE, sep = ' ')
#lab_dictionary <- read.csv('../data/eCRF_to_LOINC_forReview2.0.csv') 
lab_dictionary <- read.csv('../../laboratory_values/data/eCRF_to_TCH_forReview_copy.csv') %>% filter(!is.na(concept_cd))
ecrf <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv') %>%
  filter(Form.Name == 'laboratory_values') %>%
  select(variableName = Variable...Field.Name, Section.Header, formName = Form.Name) %>%
  mutate(Section.Header = ifelse(Section.Header == '', NA, Section.Header)) %>%
  tidyr::fill(Section.Header) 

lab_dictionary <- right_join(lab_dictionary, ecrf) 

```


Import the lab data from MISC patients. Instead of importing a subset, we import the whole dataset (ag440_music_labs_nov21) and here we filter by the concepts that are in our lab_dictionary and we rename the admission and discharge date colum.  

Here we specify for which patients we want to run the pipeline for.
```{r}
#music_patients_admission_after_nov21 <- read.csv("../../dataRefresh/local_ref/music_patients_feb2022_manually_updated.csv")

#redcap <- read.csv("../../laboratory_values/local_ref/redcap_output_laboratory_values_updated.csv") %>%
#  pull( record_id ) %>%
#  unique()

#newPatients <- music_patients_admission_after_nov21 %>%
#  filter( ! MUSIC.ID %in% redcap )


# now we have our list of new poatient IDs we want to run the pipeline on.

newPatients <- read.csv('../local_ref/texasChildren/misc_patient.csv') %>% pull(MUSIC.Patient.ID)

```


```{r}
#all_labs <- read.csv('../../dataRefresh/local_ref/AG440_MUSIC_LABS_FEB22')
all_labs <- read.csv('../local_ref/texasChildren/misc_lab_observations.csv') %>%
  #rename(CONCEPT_CD = Concept.code, 
  #       START_DATE = Date,
  #       TVAL_CHAR = Categorical.Value,
  #       NVAL_NUM = Numeric.Value,
  #       UNITS_CD = Lab.Units) %>%
  left_join(read.csv('../local_ref/texasChildren/misc_patient.csv') %>%
    mutate(MUSIC.Patient.ID = as.character(MUSIC.Patient.ID))%>%
    rename(ADMISSIONDATE = Admission.Date,
           DISCHARGE_DATE = Discharge.Date, 
           patient_num = MUSIC.Patient.ID)) %>% 
    filter(inout_cd == 'INPATIENT')

#misc_mapping <- read.csv("../../dataRefresh/local_ref/MUSIC_ID_MAPPING") %>%
#  rename(PATIENT_NUM = PAT_NUM) %>%
#  filter( MUSIC_ID %in% newPatients$MUSIC.ID )

subset_df <- all_labs %>%
  #filter( concept_cd %in% lab_dictionary$LOINC.CODE ) %>%
  rename('ADMISSION_DATE' = ADMISSIONDATE,
         'DISCHARGE_DATE' = DISCHARGE_DATE) #%>%
  #filter(PATIENT_NUM %in% misc_mapping$PATIENT_NUM) #####

# we flter the subset df to only include the patients of interest
```


Import the follow-up visit dates. The MUSIC_DATA_2021-11-23_0940.csv is the file extracted from REDCap of the data previously entered by BCH experts. From there we will extract the specific follow-up dates for the MUSIC study for the MISC patients. The variable name that contains the date is "visit_date", extracted from the dictionary. The record id is the internal MUSIC identifier, and the redcap_event_name allows us to know which is the specific follow-up period the date is referring to. Because the ID are MUSIC Ids we also need to load the music_mapping_id. 
```{r}
#visit_dates  <- read.csv("../../dataRefresh/local_ref/MUSIC_DATA_2022-03-10_1345.csv")

#visit_dates <- visit_dates %>%
#  filter( visit_date != "" ) %>%
#  select( record_id, redcap_event_name, visit_date )

### check counts: for how many patients we have all the follow-up dates (2 weeks, 6 weeks and 6 months?)
### in this case only for 12 of the patients all the information has been entered
#counts <- visit_dates %>%
#  group_by( record_id ) %>%
#  mutate( dates_available = n( ) ) %>%
#  select( record_id, dates_available ) %>%
#  unique()
#summary(as.factor( counts$dates_available))

### let's put together in a table all the information, combining the patient timeline with first admission and discharge date with the rest of follow-up dates when available
#misc_followUp <- visit_dates %>%
#  spread( redcap_event_name, visit_date ) %>%
#  left_join( misc_mapping, by = c("record_id"="MUSIC_ID") ) %>%
#  select( PATIENT_NUM, two_weeks = "2_weeks_postdischa_arm_1", 
#          six_weeks = "6_weeks_postdischa_arm_1", six_months = "6_months_postdisch_arm_1")

#head(misc_followUp)
```


Data transformation
- split `START_DATE` field into `date` and `time`
- perform a left join to combine extracted data with variables in lab dictionary
- reformat `ADMISSION_DATE` and `DISCHARGE_DATE` as dates
- calculate `days_after_admission` and `days_after_discharge`
And we join this information with the follow-up information. 
```{r}
subset_df$date <- sapply(strsplit( as.character(subset_df$start_date), " "), '[', 1)
subset_df$time <- sapply(strsplit( as.character(subset_df$start_date), " "), '[', 4)

finalTable <- left_join( subset_df, lab_dictionary, by = c('concept_cd' = 'concept_descripiton'))

finalTable <- finalTable %>%
  mutate(date = dmy(date),
         ADMISSION_DATE = mdy(ADMISSION_DATE),
         DISCHARGE_DATE = mdy(DISCHARGE_DATE),
         days_after_admission = date - ADMISSION_DATE,
         days_after_discharge = date - DISCHARGE_DATE) #%>%
  #left_join( misc_followUp )

finalTable <- finalTable %>% filter(!is.na(variableName))



```

#### Check the quality of the follow up dates

```{r}
##### SKIP FOR TCH #####
#qc_dates <- finalTable %>%
#  mutate( two_weeks_daysDiff = abs( ymd( two_weeks) - ymd( DISCHARGE_DATE)), 
#          six_weeks_daysDiff = abs( ymd( six_weeks) - ymd( DISCHARGE_DATE)), 
#          six_months_daysDiff = abs( ymd( six_months) - ymd( DISCHARGE_DATE)), 
#          check_two_weeks = ifelse( two_weeks_daysDiff > 21 | two_weeks_daysDiff < 7, "check", "ok"), 
#          check_six_weeks = ifelse( six_weeks_daysDiff > 49| six_weeks_daysDiff < 35, "check", "ok") , 
#          check_six_months = ifelse( six_months_daysDiff > 214 | six_months_daysDiff < 150, "check", "ok")) %>%
#  select( PATIENT_NUM, ADMISSION_DATE, DISCHARGE_DATE, two_weeks, six_weeks, 
#          six_months, two_weeks_daysDiff, six_weeks_daysDiff, six_months_daysDiff, 
#          check_two_weeks, check_six_weeks, check_six_months) %>%
#  unique()
```



Extract `value` based on type of variable
- focus on `_date`, `_value`, `_unit`, and `_obtained` variables (identified as most prevalent in `data_dictionary_extraction.Rmd`)
  - choices accepted for `obtained` variables: 1, 0
  - choices accepted for `value` variables: no criteria. if NVAL_NUM is not NA, then NVAL_NUM. else TVAL_CHAR
  - choices accepted for `unit` variables: no criteria. use UNITS_CD column
  - choices accepted for `date` variables: use date column in mdy format
```{r}
# Create function to extract value
getLabValue <- function(row) {
  
  # for obtained variables, if the entry is present, the 'value' is automatically 1 except for:
  # APTT/PTT that choices are 2, APTT | 3, PTT | 0, Neither	
  if (row['vartype'] == 'obtained' ) {
    if( row['variableName'] == 'aptt_ptt_obtained'){
      value = '3' #in BCH PTT is the one estimated
    }else{
      value = '1'
    }
    
  # Variables with unit listed below the valaue field will be filled with ‘-88’
  # WBC and platelets 1, x10^3/uL | 2, 10^9/L | 3, Other --> for WBC and platelets is x10^3/uL			
  # triglycerid 1, mg/dL | 2, mmol/L | 3, Other	1	 --> for BCH the values mg/dL		
  # CRP 1, mg/L | 2, mg/dL | 3, Other --> for BCH the value is mg/dL
  # IgE: Expected unit value: 'mg/dL'; current unit value: 'unit/mL'
    
  } else if (row['vartype'] == 'unit') {
    if( row['variableName'] == 'crp_unit'){
      value = '2'
    }else if( row['variableName'] == 'triglycerid_unit'){
      value = '1'
    }else if( row['variableName'] == 'wbc_unit'){
      value = '1'
    }else if( row['variableName'] == 'platelet_unit'){
      value = '1'
    }else if (row['variableName'] == 'ige_unit'){
      value = 'unit/mL'
    } else {
      value = '-88'
    }
    
  # for date variables, use the date column
  # redcap requires the following format: 
  # (NOTE: Dates must be imported here only in M/D/Y format or Y-M-D format, regardless of the specific date format designated for this field.)

  } else if (row['vartype'] == 'date') {
    value = as.character(ymd(row['date']), format = "%m/%d/%Y")
    
  # for value variables, if NVAL_NUM is not NA, use NVAL_NUM. else use TVAL_CHAR
  # for igg and igm: return 1 Positive (detected) if >11, 3 Indeterminate if 9-11, 0 Not detected if <9
    
  } else if (row['vartype'] == 'value') {
    if(!is.na(row['nval_num'])){
      
      if(row['variableName'] %in% c('anti_sars_igg_value', 'anti_sars_igm_value')){
        if(as.numeric(row['nval_num']) > 11){
          value = '1'
        } else if(as.numeric(row['nval_num']) >= 9) {
          value = '3'
        } else {
          value = '0'
        } 
      } else {
        value = as.character(as.numeric(row['nval_num']))
      }
      
    } else {
      value = row['tval_char']
    }
  }
    
  # for other variables, mark as 'UNKNOWN' and investigate; there should not be any of these for now
  else {
    value = 'UNKNOWN'
  }
   
  return(value)
}

```

```{r}
finalTable <- finalTable %>%
  # categorize the variable types
  mutate(vartype = gsub("^.*_", "", variableName)) %>%
  # remove entries with no data
  filter(!is.na(nval_num) | tval_char != '') %>%
  # remove duplicated rows
  distinct %>%
  # filter out labs where reported value is "Not Reported" or "TNP"
  filter(!tval_char %in% c("Not Reported", "TNP"))
         
finalTable$value <- apply(finalTable, 1, FUN = getLabValue)
```

lab_values_visit
1, Admission/first obtained during MIS-C hospitalization
2, Closest to discharge during MIS-C hospitalization
3, 2 Weeks Post Discharge
4, 6 Weeks Post Discharge
6, 6 Months Post Discharge


## Assign period based on the days of diff
```{r}
 finalTable <- finalTable %>%
   mutate(period = case_when(
          days_after_admission >= -2 & days_after_admission <=  2 ~ '1',
          days_after_discharge >= -3 & days_after_discharge <= 0 ~ '2', 
          days_after_discharge >= 7 & days_after_discharge < 21 ~ '3',
          days_after_discharge >= 21 & days_after_discharge <  63 ~ '4',
          days_after_discharge >= 63 & days_after_discharge < 365 ~ '6',
          TRUE ~ 'Out of period'))


#######
# Code to deal with multiple concept codes for lymphocytes and neutrophils
# We consolidate all concepts to one code
finalTable <- finalTable %>%
  mutate(concept_cd = gsub('LAB:3775664', 'LAB:3775715', concept_cd),
         concept_cd = gsub('LAB:3775663', 'LAB:3775714', concept_cd))

#######


# select which value to report when more than one lab value for the same patient and period 
finalTable <- finalTable %>%
   filter( period != "Out of period")# %>%
#   group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
#   mutate( two_weeks_ref = if_else( !is.na( ymd(two_weeks) ), ymd(two_weeks), ymd( DISCHARGE_DATE ) + days(14) ), 
#           six_weeks_ref = if_else( !is.na( ymd(six_weeks) ), ymd(six_weeks), ymd( DISCHARGE_DATE ) + days(42) ), 
#           six_months_ref =if_else( !is.na( ymd(six_months)), ymd(six_months), ymd( DISCHARGE_DATE ) + days(182) ), 
#           daysDiff = case_when(  period == 1 ~ as.character( abs(ymd(date) - ymd(ADMISSION_DATE))),
#                                  period == 2 ~ as.character( abs(ymd(date) - ymd(DISCHARGE_DATE))),
#                                  period == 3 ~ as.character( abs(ymd(date) - ymd(two_weeks_ref))),
#                                  period == 4 ~ as.character( abs(ymd(date) - ymd(six_weeks_ref))),
#                                   period == 6 ~ as.character( abs(ymd(date) - ymd(six_months_ref))),
#                                 TRUE ~ " "), 
#           daysDiff = as.numeric( daysDiff ) )

# because we do not have the period cutoff dates.. we take the first one if there are multiple
finalTableFiltered <- finalTable %>%
 group_by( patient_num, concept_cd, period ) %>%
#  slice_min( order_by = daysDiff ) %>%
  slice_min( order_by = dmy_hm(start_date) )

finalTableFilteredCheck <- finalTableFiltered %>%
  filter( vartype == "obtained") %>%
  group_by( patient_num, concept_cd, period ) %>%
  summarise( n = n())

toCheck <- finalTableFilteredCheck[ finalTableFilteredCheck$n > 1, ]
```

redcap_repeat_instance
- Using dates to determine visit number 
- note that entries inputted at different times on the same date are NOT considered as separate events / visits
```{r}
#finalTableFiltered <- finalTableFiltered %>% 
#  group_by(PATIENT_NUM) %>%
#  mutate(datetime = ymd(date)) %>%
#  arrange(datetime) %>%
#  mutate(redcap_repeat_instance = match(datetime, sort(unique(datetime)))) %>%
#  arrange(redcap_repeat_instance)
```

Because we are choosing 1 lab value per time point (the closest to the date), we should not have more than one value for redcap_repeat_instance. 

```{r}
finalTableFiltered <- finalTable

finalTableFiltered <- finalTableFiltered %>% 
  group_by(patient_num) %>%
  mutate( lab_values_visit = period, 
    redcap_repeat_instance = lab_values_visit ) %>%
  arrange(redcap_repeat_instance)
```

Pivot table wider and make compatible with RedCap formatting
```{r}
# for cases where a patient has multiple values for a lab at the exact same date / time, just choose the first one
out <- finalTableFiltered %>%
  group_by(patient_num, formName, lab_values_visit, redcap_repeat_instance, variableName) %>%
  summarise(value = first(value)) %>%
  pivot_wider(id_cols = c(patient_num, formName, lab_values_visit, redcap_repeat_instance), 
                   names_from = variableName, 
                   values_from = value, 
                   values_fill = '')

# should we remove entries with no data?
#finalTable %>% filter(is.na(NVAL_NUM), TVAL_CHAR == '')

# map BCH ids to MUSIC IDs for REDCAP
#out <- left_join(out, misc_mapping %>% select(PATIENT_NUM, MUSIC_ID))

# Rename columns for redcap
out <- out %>%
  rename('record_id' = 'patient_num',
         'redcap_repeat_instrument' = 'formName') %>%
  mutate(redcap_event_name = "repeating_forms_arm_1") %>%
  relocate(record_id, redcap_event_name, redcap_repeat_instrument, redcap_repeat_instance, lab_values_visit) %>%
  arrange(record_id, redcap_repeat_instance, lab_values_visit) %>%
  ungroup() #%>%
  #select(-'PATIENT_NUM')
```

Fill the blank 'obtained' variables with 0, indicating that this lab was not collected for that time period.
We cannot fill with 0 the variables that are under a higher category (cytokine_obtained, immunog_obtained, compl_obtained). 
We exclude those columns of being automatically filled. 
```{r}
set_obtained_values <- function(x) {ifelse (x == '', '0', x)}


excludedColnames <-c( "tnf_alpha_obtained","il_1_obtained", "il_2_obtained", "il_2_receptor_obtained", "il_4_obtained", "il_5_obtained",
                      "il_6_obtained", "il_8_obtained", "il_10_obtained", "il_12_obtained", "il_13_obtained", 
                      "il_17_obtained", "il_19_obtained", "ifn_gamma_obtained", "igg_obtained", "iga_obtained", "igm_obtained", "ige_obtained",
                      "c3_obtained", "c4_obtained", "cd3_abs_obtained", "cd3_cd4_obtained", "cd3_cd4_abs_obtained", 
                      "cd3_obtained","cd3_cd8_obtained", "cd3_cd8_abs_obtained", "cd3_cd16_cd56_obtained", "cd19_obtained", 
                      "cd19_abs_obtained","ac4_cd8_ratio_obtained", "t_sum_obtained")
colnames_to_fill <- colnames( out )[! colnames( out ) %in% excludedColnames ]

for (col in colnames_to_fill) {
  
  if (grepl('obtained',  col)) {out[,eval(col)] = sapply(out[,eval(col)], set_obtained_values)}
  
}
  

```

Replace the BCH values for sars_cov2ntpcr_value with those expected from redcap.
Our patients have Negative, NEGATIVE, Positive, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate'

Replace the BCH values for total_anti_value with those expected from redcap. 
Our patients have values ‘Non-reactive’ and ‘Reactive’, redcap is expecting 1, Detected | 0, Not detected | 3, Indeterminate
```{r}

# # We map "Negative" to 0 (Not detected) and "Positive" to 1 (Detected)
# table(out$sars_cov2ntpcr_value)
# 
# reset_sars_cov2ntpcr_values <- function(x){
#   ifelse(x == "Negative", '0', 
#          ifelse(x == "Positive", '1', x))
# }
# 
# out <- out %>% mutate(sars_cov2ntpcr_value = reset_sars_cov2ntpcr_values(sars_cov2ntpcr_value))
# 
# # we map "Non-reactive" to 0 (Not detected) and "Reactive" to 1 (Detected)
# table(out$total_anti_value)
# 
# reset_total_anti_value <- function(x){
#   ifelse(x == "Non-reactive", '0', 
#          ifelse(x == "Reactive", '1', x))
# }
# 
# out <- out %>% mutate(total_anti_value = reset_total_anti_value(total_anti_value))
# 
# table(out$total_anti_value)



```

Adding the variables:
- cytokine_obtained (1 yes, 0 no)
- immunog_obtained (1 yes, 0 no)
- compl_obtained (1 yes, 0 no)
- other_labs ( 1 yes, 0 no)
```{r}

### If there is not a column, make a blank one
 #out$immunog_obtained <- NA
 #out$igg_obtained <- NA
 #out$igm_obtained <- NA
 #out$ige_obtained <- NA
 #out$c3_obtained <- NA
 #out$c4_obtained <- NA
 #out$cd3_obtained <- NA
 #out$cd3_abs_obtained <- NA
 #out$cd3_cd4_obtained <- NA
 #out$cd3_cd4_abs_obtained <- NA
 #out$cd3_cd8_obtained <- NA
 #out$lympho_obtained <- NA
 #out$cd3_cd8_abs_obtained <- NA
 #out$cd19_obtained <- NA
 #out$cd3_cd16_cd56_obtained <- NA
 #out$ac4_cd8_ratio_obtained <- NA
 #out$tnf_alpha_obtained <- NA
 #out$il_1_obtained <- NA
 out$il_2_obtained <- NA
 #out$il_2_receptor_obtained <- NA
 #out$il_4_obtained <- NA
 #out$il_5_obtained <- NA
 #out$il_6_obtained <- NA
 #out$il_8_obtained <- NA
 #out$il_10_obtained <- NA
 out$il_12_obtained <- NA
 out$il_13_obtained <- NA
 out$il_17_obtained <- NA
 #out$ifn_gamma_obtained <- NA
 #out$igg_obtained <- NA
 #out$iga_obtained <- NA
 #out$cd3_cd8_obtained <- NA

out <- out %>% 
  group_by( record_id , redcap_repeat_instance ) %>%
  mutate( cytokine_obtained = ifelse( tnf_alpha_obtained == 1 | il_1_obtained == 1 | il_2_obtained == 1 |
                                        il_2_receptor_obtained == 1 | il_4_obtained == 1 | il_5_obtained == 1 | 
                                        il_6_obtained == 1 | il_8_obtained == 1 | il_10_obtained == 1 | 
                                        il_12_obtained == 1 | il_13_obtained == 1 | il_17_obtained == 1  | 
                                        ifn_gamma_obtained == 1, 1, 0), 
          immunog_obtained = ifelse( igg_obtained == 1 | iga_obtained == 1 | igm_obtained == 1|  ige_obtained == 1, 1, 0),
          compl_obtained = ifelse( c3_obtained == 1 | c4_obtained == 1, 1, 0), 
          lympho_obtained = ifelse( cd3_obtained == 1 | cd3_abs_obtained == 1 | cd3_cd4_obtained == 1 | cd3_cd4_abs_obtained == 1 |
                                      cd3_cd8_obtained == 1 | cd3_cd8_abs_obtained == 1 | cd3_cd16_cd56_obtained == 1 | 
                                      cd19_obtained == 1 | ac4_cd8_ratio_obtained == 1, 1, 0), 
          other_labs = 0
)
```

Fill the blank 'obtained' variables with 0, indicating that this lab was not collected for that time period.
Do this for that are under a higher category (cytokine_obtained, immunog_obtained, compl_obtained) IF the category obtained value is TRUE. 
```{r}


out <- out %>% 
  mutate(tnf_alpha_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(tnf_alpha_obtained), ''),
         il_1_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_1_obtained), ''),
         il_2_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_2_obtained), ''),
         il_2_receptor_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_2_receptor_obtained), ''),
         il_4_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_4_obtained), ''),
         il_5_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_5_obtained), ''),
         il_6_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_6_obtained), ''),
         il_8_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_8_obtained), ''),
         il_10_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_10_obtained), ''),
         il_12_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_12_obtained), ''),
         il_13_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_13_obtained), ''),
         il_17_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(il_17_obtained), ''),
         ifn_gamma_obtained = ifelse(cytokine_obtained == 1, set_obtained_values(ifn_gamma_obtained), ''),
         igg_obtained = ifelse(immunog_obtained == 1, set_obtained_values(igg_obtained), ''),
         iga_obtained = ifelse(immunog_obtained == 1, set_obtained_values(iga_obtained), ''),
         igm_obtained = ifelse(immunog_obtained == 1, set_obtained_values(igm_obtained), ''),
         ige_obtained = ifelse(immunog_obtained == 1, set_obtained_values(ige_obtained), ''),
         c3_obtained = ifelse(compl_obtained == 1, set_obtained_values(c3_obtained), ''),
         c4_obtained = ifelse(compl_obtained == 1, set_obtained_values(c4_obtained), ''),
         cd3_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_obtained), ''),
         cd3_abs_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_abs_obtained), ''),
         cd3_cd4_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_cd4_obtained), ''),
         cd3_cd4_abs_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_cd4_abs_obtained), ''),
         cd3_cd8_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_cd8_obtained), ''),
         cd3_cd8_abs_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_cd8_abs_obtained), ''),
         cd3_cd16_cd56_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd3_cd16_cd56_obtained), ''),
         cd19_obtained = ifelse(lympho_obtained == 1, set_obtained_values(cd19_obtained), ''),
         ac4_cd8_ratio_obtained = ifelse(lympho_obtained == 1, set_obtained_values(ac4_cd8_ratio_obtained), ''))
  

```

There are some variables where we have a concept code for them in our mapping, but they are not present in any TCH MUSIC patients.
We add them here with the value for obtained being 0 IF there are other labs in their panel... otherwise leave blank
- troponin I
- ah50
- ch50
- nt-probnp
- cxcl9
- 
```{r}

out <- out %>% 
  mutate(
    #troponin_l_obtained  = 0,
         ah50_obtained  = ifelse(compl_obtained == 1, yes = '0', no = ''), 
         ch50_obtained = ifelse(compl_obtained == 1, yes = '0', no = ''),
    #     nt_probnp_obtained = 0,
    #    cxcl9_obtained = ifelse(cytokine_obtained == 1, yes = '0', no = ''),
         sars_cov2_antig_obtained = '0')


```

Create a report for BCH of which patients / values require manual review
```{r}
flagged <- data.frame('record_id' = '', 
                      'variable' = '',
                      'visit' = '',
                      'value' = '')[-1,]

report_manual_review <- function(r,c){}

value_vars <- grep('_value$', colnames(out))
unexpected_values = c('SEE IL12', 'See Image', 'Not Reported', 'Footnote', 'SEE IL17')

# for each row in out
for (r in c(1:nrow(out))){
  
  record_id <- as.character(out$record_id[r])
  visit <- out$lab_values_visit[r]

  # for each value variable
  for (c in value_vars) {
    
    variable <- colnames(out)[c]

    # if value is in our list of weird things
    if (out[r,c] %in% unexpected_values) {
    
      # if value is specifically "Not Reported"
      ####### NOTE THIS SHOULD NOT HAPPEN ANY MORE
      #if(out[r,c] == "Not Reported"){
        
        # clear all related values and set obtained to 0
      #  prefix <- gsub('_value', '', colnames(out)[c])
      #  out[r,which(grepl(prefix, colnames(out)))] <- ''
      #  out[r,which(grepl(paste0(prefix, '_obtained'), colnames(out)))] <- '0'
        
      # else add record, variable, visit, value, to flagged df and just delete weird thing from out
      #} else {
        
        flagged <- flagged %>% 
        add_row('record_id' = record_id,
               'variable' = variable,
               'visit' = visit,
               'value' = out[r,c] %>% pull)
      
        out[r,c] <- ''
      #}
      

    }
    
  }
}



```


Data formatting based on errors / warnings from BCH upload to REDCap
```{r}
#out <- out %>% 
#  mutate(il_1_value = ifelse(il_1_value == "SEE IL12", '', il_1_value),
#         il_10_value = ifelse(il_10_value == "SEE IL12", '', il_10_value),
#         il_13_value = ifelse(il_13_value == "SEE IL12", '', il_13_value),
#         il_2_value = ifelse(il_2_value == "SEE IL12", '', il_2_value),
#         il_5_value = ifelse(il_5_value == "SEE IL12", '', il_5_value),
#         il_8_value = ifelse(il_8_value == "SEE IL12", '', il_8_value),
#         il_12_value = ifelse(il_12_value == "See Image", '', il_12_value),
#         ast_sgot_value = ifelse(ast_sgot_value == "Not Reported", '', ast_sgot_value),
#         ast_sgpt_value = ifelse(ast_sgpt_value == "Not Reported", '', ast_sgpt_value),
#         il_2_receptor_value = ifelse(il_2_receptor_value == "SEE IL12", '', il_2_receptor_value),
#         il_4_value = ifelse(il_4_value == "SEE IL12", '', il_4_value),
#         il_6_value = ifelse(il_6_value == "SEE IL12", '', il_6_value),
#         troponin_t_value = ifelse(troponin_t_value == "Not Reported", '', troponin_t_value),
#         direct_bilir_value = ifelse(direct_bilir_value == "Not Reported", '', direct_bilir_value),
#         ige_value = ifelse(ige_value == "Footnote", '', ige_value),
#         lymphocyt_value = ifelse(lymphocyt_value == "Footnote", '', lymphocyt_value),
#         neutrophil_value = ifelse(neutrophil_value == "Footnote", '', neutrophil_value),
#         ifn_gamma_value = ifelse(ifn_gamma_value == "SEE IL17", '', ifn_gamma_value),
#         il_1_value = ifelse(il_1_value == "SEE IL17", '', il_1_value),
#         il_10_value = ifelse(il_10_value == "SEE IL17", '', il_10_value),
#         il_12_value = ifelse(il_12_value == "SEE IL17", '', il_12_value),
#         il_13_value = ifelse(il_13_value == "SEE IL17", '', il_13_value),
#         il_17_value = ifelse(il_17_value == "SEE IL17", '', il_17_value),
#         #il_19_value = ifelse(il_19_value == "SEE IL17", '', il_19_value),
#         il_2_receptor_value = ifelse(il_2_receptor_value == "SEE IL17", '', il_2_receptor_value),
#         il_2_value = ifelse(il_2_value == "SEE IL17", '', il_2_value),
#         il_4_value = ifelse(il_4_value == "SEE IL17", '', il_4_value),
#         il_5_value = ifelse(il_5_value == "SEE IL17", '', il_5_value),
#         il_6_value = ifelse(il_6_value == "SEE IL17", '', il_6_value),
#         il_8_value = ifelse(il_8_value == "SEE IL17", '', il_8_value),
#         tnf_alpha_value = ifelse(tnf_alpha_value == "SEE IL17", '', tnf_alpha_value),
#         ggtp_value = ifelse(ggtp_value == "Not Reported", '', ggtp_value),
#         ifn_gamma_value = ifelse(ifn_gamma_value == "SEE IL12", '', ifn_gamma_value),
#         il_12_value = ifelse(il_12_value == "SEE IL12", '', il_12_value),
#         il_17_value = ifelse(il_17_value == "SEE IL12", '', il_17_value),
         #il_19_value = ifelse(il_19_value == "SEE IL12", '', il_19_value),
#         tnf_alpha_value = ifelse(tnf_alpha_value == "SEE IL12", '', tnf_alpha_value),
#         bicarbonate_value = ifelse(bicarbonate_value == "Not Reported", '', bicarbonate_value),
#         hematocrit_value = ifelse(hematocrit_value == "Footnote", '', hematocrit_value),
#         hemoglobin_value = ifelse(hemoglobin_value == "Footnote", '', hemoglobin_value),
#         platelets_value = ifelse(platelets_value == "Footnote", '', platelets_value),
#         wbc_value = ifelse(wbc_value == "Footnote", '', wbc_value),
#         ferritin_value = ifelse(ferritin_value == "Not Reported", '', ferritin_value),
#         potassium_value = ifelse(potassium_value == "Not Reported", '', potassium_value))
  
# write output with patient IDs included
write.csv(out, '../local_ref/lab_output_texas.csv', row.names = FALSE)
#write.csv(out, '../local_ref/redcap_output_laboratory_values_updated.csv', row.names = FALSE)
#write.csv(out %>% filter(record_id %in% c(120002, 120122, 120018, 120006)), '../local_ref/TEST_redcap_output_laboratory_values_updated.csv', row.names = FALSE)
#write.csv(flagged, '../local_ref/flagged_labs_for_manual_review.csv', row.names = FALSE)
```













