---
title: "Generate Potential eCRF to LOINC Mapping"
author: "Simran Makwana"
date: "6/11/2021"
output: html_document
---

# Load Packages
```{r}
library(dplyr)
library(lubridate)

```

# Load UMLS database subset (all LOINCs which are clinical attributes)
The following SQL was used to generate the subset:
create table ag440_loinc_clinicalAttributes as
select conso.cui, conso.lat, conso.sab, conso.tty, conso.code, conso.str, sty.sty 
from mrconso conso, mrsty sty 
where conso.cui = sty.cui and
conso.sab = 'LNC' and conso.tty = 'LN' and sty.sty = 'Clinical Attribute';

```{r}
umls_subset <- read.delim("../local_ref/loinc_clinicalAttributes.dsv")

```

# Load the manually generated keywords
The full list of keywords can be viewed in the GitHub repository at this link.
Set the 'cutOff' value. Keyword queries which exceed the cutoff will have the additional 'Ser/Plas' and 'Pt:Bld' filters applied.
```{r}
lab_keywords <- read.delim("../data/lab_keywords.tsv" )
lab_keywords <- lab_keywords %>%
  dplyr::filter( Position %in% c('beginning', 'anywhere'))
cutOff <- 10
```

# Extract the list of potential LOINC codes
Search through the UMLS subset for the keywords associated with the eCRF terms.
```{r}

for( i in 1:nrow(lab_keywords)){
  position <- lab_keywords$Position[i]
  if( position == "anywhere"){
    generalQ<- umls_subset %>% 
      dplyr::filter( grepl( lab_keywords$Keyword[i], STR))
    }else if(position == "beginning"){
      generalQ<- umls_subset %>% 
        dplyr::filter( grepl( paste0("^", lab_keywords$Keyword[i]), STR))
  }
  
  narrowQ <- generalQ %>% 
    dplyr::filter( grepl( paste(c("*Ser/Plas:Qn", "*Pt:Bld:Qn"), collapse = "|"), STR))
  
  if(i==1){
    if( nrow(narrowQ) == 0 ){
      output <- generalQ
    }else if( nrow( generalQ ) > cutOff ){
      output <- narrowQ
    }else if( nrow( generalQ) > 0 ){
      output <- generalQ
    }else{
      print( paste0("No entry found for keyword: ", lab_keywords$Keyword[i]))
    }
    print( paste0( nrow( output), " LOINC terms found for eCRF Term: ", lab_keywords$Section.Header[i], " with keyword: ", lab_keywords$Keyword[i]))
    output$Section.Header <- lab_keywords$Section.Header[i]
  }else{
    if( nrow(narrowQ) == 0 ){
      int_output <- generalQ
    }else if( nrow( generalQ ) > cutOff ){
      int_output <- narrowQ
    }else if( nrow( generalQ) > 0 ){
      int_output <- generalQ
    }else{
      print( paste0("No entry found for keyword: ", lab_keywords$Keyword[i]))
    }
    print( paste0( nrow( int_output), " LOINC terms found for eCRF Term: ", lab_keywords$Section.Header[i], " with keyword: ", lab_keywords$Keyword[i]))
    int_output$Section.Header <- lab_keywords$Section.Header[i]
    output <- rbind( output, int_output)
  }
}

output <- unique(output)
ranked_list <- output %>% group_by(Section.Header) %>% summarise(n = n_distinct(CODE)) %>% arrange(desc(n)) # testing


```
# Save list of potential LOINC codes for clinician check

```{r}
out <- output %>%
  mutate(webpage = paste0("https://loinc.org/",CODE,"/"),
         clinical_review = '') %>%
  rename(`LOINC CODE` = CODE,
         `LOINC DESCRIPTION` = STR,
         `eCRF TERM DESCRIPTION` = Section.Header,
         `CLINICIAN REVIEW: DO THE DESCRIPTIONS MATCH?` = clinical_review,
         `LOINC WEBPAGE - EXTENDED DESCRIPTION` = webpage)

write.csv(out, '../data/eCRF_to_LOINC_forReview.csv', row.names = FALSE)

```

# Site - Specific Steps

## BCH

### Which of the identified potential LOINC codes are present in the site-specific concept mapping?
```{r}
bch_loinc_map <- read.csv("../data/BCH_Lab_Loinc_cd_Map_data.csv")

mappingToBch <- bch_loinc_map %>%
  dplyr::filter(LOINC_LAB_CODE %in% output$CODE)

```

### Of those LOINC codes which map to the internal concepts, which many LOINC codes are present in the MIS-C patients at this site?
```{r}
observationFactsLabs <- read.delim("../local_ref/music_labs_observation_withUnits.dsv")
observationFactsLabs$date <- sapply(strsplit( as.character(observationFactsLabs$START_DATE), "[ ]"), '[', 1)
colnames( observationFactsLabs ) <- tolower( colnames( observationFactsLabs ) )

labs_in_misc <- observationFactsLabs %>%
  dplyr::filter( concept_cd %in% mappingToBch$BCH_LAB_CODE )

labs_in_misc$units_cd <- gsub( "MMOL/L", "mmol/L", labs_in_misc$units_cd)
labs_in_misc$units_cd <- gsub( "MG/DL", "mg/dL", labs_in_misc$units_cd)
labs_in_misc$units_cd <- gsub( "G/DL", "g/dL", labs_in_misc$units_cd)

check_summary <- labs_in_misc %>%
  dplyr::group_by(concept_cd, units_cd) %>%
  dplyr::summarise( min_val = min(nval_num, na.rm = TRUE), 
                    max_val = max(nval_num, na.rm = TRUE), 
                    min_date = min( dmy( date ) ), 
                    max_date = max( dmy( date ) ), 
                    distinct_patients = length(unique(patient_num)), 
                    distinct_observations = length(patient_num)) %>%
  dplyr::select( concept_cd, min_date, max_date, 
                 distinct_patients, distinct_observations, 
                 min_val, max_val, units_cd)

colnames(mappingToBch) <- c("concept_cd", "CODE")
check_summary <- left_join( check_summary, mappingToBch )
check_summary <- left_join( check_summary, output )

check_summary <- check_summary %>% 
  dplyr::filter( units_cd != 'NOT DEFINED IN SOURCE')


```