---
title: "Generate Potential eCRF to LOINC Mapping: refinement after checking UTAH mapping"
author: "Simran Makwana & Alba Gutierrez"
date: "5/13/2022"
output: html_document
---

# Load Packages
```{r}
library(dplyr)
library(lubridate)
library(readxl)
```

# Load UMLS database subset (all LOINCs which are clinical attributes)
The following SQL was used to generate the subset:
create table ag440_loinc_clinicalAttributes as
select conso.cui, conso.lat, conso.sab, conso.tty, conso.code, conso.str, sty.sty 
from mrconso conso, mrsty sty 
where conso.cui = sty.cui and
conso.sab = 'LNC' and conso.tty = 'LN' and sty.sty = 'Clinical Attribute';

```{r}
umls_subset <- read.delim("../local_ref/loinc_clinicalAttributes.dsv")
```

# Load the manually generated keywords
The full list of keywords can be viewed in the GitHub repository at this link.
Same keywords have been modified compared to the lab_keywords first version. 
- Creatinine keyword was before Creatine:, now is creatinine:
- for uera, we have make it more general, from Urea: to Urea.
- LDH did not have any keyword, added Lactate dehydrogenase:
Set the 'cutOff' value. Keyword queries which exceed the cutoff will have the additional 'Ser/Plas' and 'Pt:Bld' filters applied.
```{r}
lab_keywords <- read.delim("../data/lab_keywords2.0.tsv" )
lab_keywords <- lab_keywords %>%
  dplyr::filter( Position %in% c('beginning', 'anywhere'))
cutOff <- 10
```

# Extract the list of potential LOINC codes
Search through the UMLS subset for the keywords associated with the eCRF terms.
```{r}

for( i in 1:nrow(lab_keywords)){
  position <- lab_keywords$Position[i]
  if( position == "anywhere"){
    generalQ<- umls_subset %>% 
      dplyr::filter( grepl( lab_keywords$Keyword[i], STR))
    }else if(position == "beginning"){
      generalQ<- umls_subset %>% 
        dplyr::filter( grepl( paste0("^", lab_keywords$Keyword[i]), STR))
  }
  
  narrowQ <- generalQ %>% 
    dplyr::filter( grepl( paste(c("*Ser/Plas:Qn", "*Pt:Bld:Qn"), collapse = "|"), STR))
  
  if(i==1){
    if( nrow(narrowQ) == 0 ){
      output <- generalQ
    }else if( nrow( generalQ ) > cutOff ){
      output <- narrowQ
    }else if( nrow( generalQ) > 0 ){
      output <- generalQ
    }else{
      print( paste0("No entry found for keyword: ", lab_keywords$Keyword[i]))
    }
    print( paste0( nrow( output), " LOINC terms found for eCRF Term: ", lab_keywords$Section.Header[i], " with keyword: ", lab_keywords$Keyword[i]))
    output$Section.Header <- lab_keywords$Section.Header[i]
  }else{
    if( nrow(narrowQ) == 0 ){
      int_output <- generalQ
    }else if( nrow( generalQ ) > cutOff ){
      int_output <- narrowQ
    }else if( nrow( generalQ) > 0 ){
      int_output <- generalQ
    }else{
      print( paste0("No entry found for keyword: ", lab_keywords$Keyword[i]))
    }
    print( paste0( nrow( int_output), " LOINC terms found for eCRF Term: ", lab_keywords$Section.Header[i], " with keyword: ", lab_keywords$Keyword[i]))
    int_output$Section.Header <- lab_keywords$Section.Header[i]
    output <- rbind( output, int_output)
  }
}

output <- unique(output)
ranked_list <- output %>% group_by(Section.Header) %>% summarise(n = n_distinct(CODE)) %>% arrange(desc(n)) # testing


```
# Save list of potential LOINC codes for clinician check

```{r}
out <- output %>%
  mutate(webpage = paste0("https://loinc.org/",CODE,"/"),
         clinical_review = '') %>%
  rename(`LOINC CODE` = CODE,
         `LOINC DESCRIPTION` = STR,
         `eCRF TERM DESCRIPTION` = Section.Header,
         `CLINICIAN REVIEW: DO THE DESCRIPTIONS MATCH?` = clinical_review,
         `LOINC WEBPAGE - EXTENDED DESCRIPTION` = webpage)

write.csv(out, '../data/eCRF_to_LOINC_forReview2.0.csv', row.names = FALSE)

```

# Site - Specific Steps

## UTAH

## Add manually some loincs
```{r}
ptt <- c("C0798152", "ENG", "LNC", "LN", "14979-9", "Coagulation surface induced:Time:Pt:PPP:Qn:Coag", " Clinical Attribute", "APTT or PTT (seconds)") # 61 music patients
pt <- c("C0482694", "ENG", "LNC", "LN", "5902-2", "Coagulation tissue factor induced:Time:Pt:PPP:Qn:Coag", "Clinical Attribute", "PT (seconds)") # 60 music patients
lactate <- c("C1147817", "ENG", "LNC", "LN", "32133-1", "Lactate:SCnc:Pt:PlasV:Qn", "Clinical Attribute", "Lactate (mmol/L)") # 32 patients
dimer <- c("C1953449", "ENG", "LNC", "LN", "48065-7", "Fibrin D-dimer FEU:MCnc:Pt:PPP:Qn", "Clinical Attribute","D-dimer (mcg/mL FEU)") #66 patients
directBilirrubin <- c("C0364101", "ENG", "LNC", "LN", "1968-7", "Bilirubin.glucuronidated+Bilirubin.albumin bound:MCnc:Pt:Ser/Plas:Qn", "Clinical Attribute", "Direct bilirubin (mg/dL)") #(only present in 3 music patients)
totalBilirrubin <- c("C0364108", "ENG", "LNC", "LN", "1975-2", "Bilirubin:MCnc:Pt:Ser/Plas:Qn", "Clinical Attribute", "Total bilirubin (mg/dL)") #66 music patients
output <- rbind( output, ptt, pt, lactate, dimer, directBilirrubin, totalBilirrubin)
```


### Which of the identified potential LOINC codes are present in the site-specific concept mapping?
```{r}
utah_loinc_map <- read_xlsx("../local_ref/MUSIC lab tests for potential mapping.xlsx")

mappingToUtah <- utah_loinc_map %>%
  dplyr::filter(LOINC_CD %in% output$CODE)

labecrfCategoryInUtah <- output %>%
  dplyr::filter( CODE %in% utah_loinc_map$LOINC_CD) %>%
  select( Section.Header) %>%
  unique()

tocheck <- output %>%
  dplyr::filter(! Section.Header %in% labecrfCategoryInUtah$Section.Header) %>%
  select( Section.Header) %>%
  unique()
```


