---
title: 'ETL: Medications During Hospitalization eCRF: data refresh'
author: "Simran Makwana & Alba Gutierrez"
date: "2/16/2022"
output: html_document
---

# This notebook contains the following steps:
1.Load data extract for the concepts of interest
2. Determine expected values for REDCap 
3. Calculate medcur1_current column
4. Transform and format
5. Select only the new patients added compared to the latest refresh

# 0. Libraries
```{r}
library(tidyverse)
library(lubridate)
```

# 1. Load data extract for the codes of interest
Use the same query as medications_before_and_after -- see that pipeline for details
```{r}

df <- read.delim('../local_ref/ag440_music_meds_feb22.dsv')

# if the code is formatted like so: '2.00', convert to '02.00'
fix_code_formatting <- function(x) {
  ifelse(test = !grepl("[[:digit:]]{2}.", x), yes = (paste0('0', x)), no = (x))
}

datadict <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv') %>%
  filter(Form.Name == 'additional_medications_during_hospitalization') %>%
  select(Variable...Field.Name, Choices..Calculations..OR.Slider.Labels) %>%
  separate(Variable...Field.Name, c('a', 'b', 'c'), sep = '_') %>%
  filter(!b %in% c('na', 'type', 'name', 'start', 'ae')) %>%
  mutate(a = str_replace(a, '[[:digit:]]+', ''),
         Variable...Field.Name = paste0(a, 'X', '_', b, '_', c)) %>%
  unique() %>% 
  separate_rows(Choices..Calculations..OR.Slider.Labels, sep = '\\|') %>%
  mutate(ECRF_CODE = str_replace(Choices..Calculations..OR.Slider.Labels, ',.*$', ''),
         ECRF_CODE = str_trim(ECRF_CODE),
         ECRF_CODE = fix_code_formatting(ECRF_CODE)) %>%
  select(-Choices..Calculations..OR.Slider.Labels)

bch_med_concept_summary <- read_csv('../../medications_before_and_after/data/bch_med_concept_summary_toReview.csv') %>%
  filter(RXNORM_DESCS != 'greater than 5 RXNORM codes mapped to BCH code' | (CONCEPT_CD == 'ADMINMED:2797919' & ECRF_CODE == '07.08'))

bch_med_concept_summary <- left_join(bch_med_concept_summary, datadict, by = 'ECRF_CODE')

df <- left_join(df, bch_med_concept_summary %>% 
                  select(ECRF_CODE, CONCEPT_CD, Variable...Field.Name), by = 'CONCEPT_CD') 

# NAs that appear are not present in our mapping / are not requested by the ECRF
# we assign them the 99.99 code for 'other'
df <- df %>%
  mutate(ECRF_CODE = ifelse(is.na( ECRF_CODE ),
                            '99.99',
                            ECRF_CODE))
```


```{r}

# we have some concepts which we want to remove, even from the 'other' mappings
concepts_toremove <- read.csv('../../medications_before_and_after/data/bch_concepts_to_remove.csv') %>% pull(CONCEPT_CD)


# for 'other' medications, load in bch concept_dimension table for medications.
# we onlyh report the names for unknown medications.
# even though we are reporting all 'other' medications, we will remove the general sodium chloride concepts as they are not informative.
bch_concept_dim <- read.delim('../../medications_before_and_after/data/concept_dimensionMeds.dsv') %>%
  filter(!CONCEPT_CD %in% concepts_toremove) %>%
  select(CONCEPT_CD, NAME_CHAR) %>%
  # we have to manually add the concept for 'albumin human', based on clinician review with Simran Mahanta and Audery Dionne
  add_row(CONCEPT_CD = 'ADMINMED:2797321', NAME_CHAR = 'albumin human') %>%
  unique %>%
  group_by(CONCEPT_CD) %>% 
  summarise(NAME_CHAR = paste0(NAME_CHAR, collapse = ', '))



df <- df %>% 
  left_join(bch_concept_dim, by = 'CONCEPT_CD') %>%
  mutate(medhospX_name = ifelse(ECRF_CODE == '99.99',
                                tolower(NAME_CHAR),
                                '')) %>%
  filter(!is.na(medhospX_name))





```


```{r}
# which medications do we report dates for?
report_dates <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv') %>%
  filter(Form.Name == 'additional_medications_during_hospitalization', 
         Variable...Field.Name == 'medhosp1_start_dt') %>%
  pull(Branching.Logic..Show.field.only.if....)

report_dates

# we write the following regular expression to capture all of these codes:
report_dates <- '26.|28.|99.99|09.99|09.03|09.09|18.03|18.04|18.06|18.09|18.10'
```

# 2. Filter to medication values during hospitalization

```{r}
df <- df %>%
  mutate(date = dmy(sapply(strsplit( as.character(START_DATE), " "), '[', 1)),
         DISCHARGE_DATE = dmy(DISCHARGE_DATE),
         ADMISSION_DATE = dmy(ADMISSIONDATE),
         days_after_admission = date - ADMISSION_DATE,
         days_after_discharge = date - DISCHARGE_DATE) %>%
  filter(days_after_admission >= 0,
         days_after_discharge <= 0)
```

# Add additional columns
- medhospX_type: overall category of medication
- medhospX_start_dt: start date, mm-dd-yyyy *****This does not appear to be needed
- medhospX_name: name of medication, only include for 'other' medications
- med_num refers to the 'X' in the variable name. (internal)
- redcap_repeat_instance refers to the form number. you can have 10 med_nums per form.
```{r}

# only report the dates for the specified medications. for all other medications, medhosp_start_dt should be blank
# for medications where we are reporting the date, only report the 'start date', or the earliest date.
df2 <- df %>% 
  mutate(medhospX_start_dt = format(date, "%m/%d/%Y")) %>%
  group_by(PATIENT_NUM, ECRF_CODE, medhospX_name) %>%
  summarise(medhospX_start_dt = ifelse(grepl(report_dates, ECRF_CODE),
                                    min(medhospX_start_dt),
                                    ''),
            Variable...Field.Name = first(Variable...Field.Name)) %>%
  ungroup() %>%
  unique()



df3 <- df2 %>%
  group_by(PATIENT_NUM) %>%
  arrange(ECRF_CODE) %>%
  mutate(rownum = row_number(),
         med_num = ifelse(rownum %% 10 == 0, 10, rownum %% 10),
         redcap_repeat_instance = (rownum - 1) %/% 10 + 1,
         medhospX_type = ifelse(ECRF_CODE == '99.99',
                                '99.99',
                                (str_replace(str_replace(ECRF_CODE, '\\...', ''), '^0', '')))) %>%
  ungroup() %>%
  select(-rownum) 

# rename X in variable name depending on 'med_num'

df4 <- df3 %>%
  mutate(Variable...Field.Name = str_replace(Variable...Field.Name, 'X', as.character(med_num)))



```

# Pivot into final format
```{r}

# pivot variable names
out <- pivot_wider(df4,
                  id_cols = c(PATIENT_NUM, medhospX_type, medhospX_name, redcap_repeat_instance, med_num, medhospX_start_dt),   
                  names_from = Variable...Field.Name,
                  values_from = ECRF_CODE,
                  values_fill = NA) 

# pivot medhospX_start_dt
out2 <- out %>%
  mutate(medhospX_start_dt_names = str_replace('medhospX_start_dt', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_start_dt_names,
    values_from = medhospX_start_dt,
    values_fill = NA)
  

# pivot medhospX_type
out3 <- out2 %>%
  mutate(medhospX_type_names = str_replace('medhospX_type', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_type_names,
    values_from = medhospX_type,
    values_fill = NA)

# pivot medhospX_name
out4 <- out3 %>%
  mutate(medhospX_name_names = str_replace('medhospX_name', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_name_names,
    values_from = medhospX_name,
    values_fill = NA) %>%
  select(-med_num)


# map BCH ids to MUSIC IDs for REDCAP
id_mapping <- read.csv('../../local_ref/music_id_mapping.csv')
out5 <- left_join(out4, id_mapping %>% select(PATIENT_NUM, MUSIC_ID)) %>% 
  ungroup() %>% 
  select(-PATIENT_NUM) %>%
  rename(record_id = MUSIC_ID) %>%
  mutate(redcap_event_name = "day_1_arm_1",
         redcap_repeat_instrument = "additional_medications_during_hospitalization") %>%
  relocate(record_id, redcap_event_name, redcap_repeat_instrument) 
  
# collapse rows by form
out6 <- out5 %>%
  group_by(record_id, redcap_repeat_instance) %>%
  summarise_all(~first(na.omit(.)))
#out5 <- out4 %>%
#  group_by(record_id, redcap_repeat_instance) %>%
#  summarise_each(funs(first(.[!is.na(.)])))

# remove NA column introduced by 'other' variables
out7 <- out6 %>% select(-grep('NA', colnames(out6)))

# write output with patient IDs included
write.csv(out4, '../local_ref/redcap_output_medications_during_all_updated.csv', row.names = FALSE, na = '')
#write.csv(out3 %>% select(-PATIENT_NUM, -redcap_repeat_instance), '../local_ref/redcap_output_medications_during.csv', row.names = FALSE)
```

```{r}
### write test output for test patients 120002 and 120003
write.csv(out7 %>%
            filter(record_id %in% c(120006, 120002, 120095, 120123)) %>%
            mutate(record_id = ifelse(record_id == 120006,
                                      888006, 
                                      ifelse(record_id == 120002,
                                             888002,
                                             ifelse(record_id == 120095,
                                                    888095,
                                                    888123)))),
          '../local_ref/redcap_TEST_output_medications_during_with_latest_data.csv',
          row.names = FALSE, 
          na = '')


```

# Load the lates and the new output and compare both

```{r}
rm(list=ls()) #clean the environment

newMedsForRedcap <- read_csv( '../local_ref/redcap_output_medications_during_all_updated.csv')
previousMedsPushedToRedcap <- read_csv( '../../medications_during/local_ref/redcap_output_medications_during.csv')

newPatientsMeds <- newMedsForRedcap %>%
  filter(! PATIENT_NUM %in% previousMedsPushedToRedcap$PATIENT_NUM)

write.csv(newPatientsMeds, '../local_ref/redcap_output_medications_during_only_new_patients.csv', row.names = FALSE, na = '')

```

# QC
As a sanity check we compare the output for the test patients, to make sure that nothing else changed on the data for previous patients.
```{r}
rm(list=ls()) #clean the environment

newMedsForRedcap_TestPatients <- read_csv( '../local_ref/redcap_TEST_output_medications_during_with_latest_data.csv')
previousMedsPushedToRedcap_TestPatients <- read_csv( '../../medications_during/local_ref/redcap_TEST_output_medications_during.csv') %>%
  filter( record_id %in% newMedsForRedcap_TestPatients$record_id )

# check the columns
dplyr::all_equal( newMedsForRedcap_TestPatients, previousMedsPushedToRedcap_TestPatients )

#install.packages("janitor")
library(janitor)
compare_df_cols( newMedsForRedcap_TestPatients, previousMedsPushedToRedcap_TestPatients )

```



