---
title: 'Lab Values: Percentage of overwriting cases'
author: "Simran Makwana & Alba Gutierrez"
date: "03/01/2022"
output: html_document
---


```{r}
library(tidyverse)
library(lubridate)
lab_dictionary <- read.delim("../../laboratory_values/data/labDictionary_updated.txt", header = TRUE, sep = ' ')
```


# Run the pipeline until the finalTableFiltered is created with the latest data update and save it as an RData file
Import the lab data from MISC patients. Instead of importing a subset, we import the whole dataset (ag440_music_labs_nov21) and here we filter by the concepts that are in our lab_dictionary and we rename the admission and discharge date colum.  
```{r}
all_labs <- read.csv('../../local_ref/AG440_MUSIC_MEDS_FEB22.csv')
subset_df <- all_labs %>%
  filter( CONCEPT_CD %in% lab_dictionary$concept_cd ) %>%
  rename('ADMISSION_DATE' = ADMISSIONDATE,
         'DISCHARGE_DATE' = DISCHARGE_DATE)
```

Import the follow-up visit dates. The MUSIC_DATA_XXXX.csv is the file extracted from REDCap of the data previously entered by BCH experts. From there we will extract the specific follow-up dates for the MUSIC study for the MISC patients. The variable name that contains the date is "visit_date", extracted from the dictionary. The record id is the internal MUSIC identifier, and the redcap_event_name allows us to know which is the specific follow-up period the date is referring to. Because the ID are MUSIC Ids we also need to load the music_mapping_id. 

To create the music_id_mapping file we create this table in SQL:
create table music_id_mapping_feb22 as 
select m.MRN, m.MUSIC_ID, p.PAT_NUM from MUSIC_PATIENTLIST_FEB22 m, MRN_PATUUID_PATNUM p where m.MRN = p.MRN; 

```{r}
misc_mapping <- read.csv("../../local_ref/music_id_mapping.csv")
visit_dates  <- read.csv("../../local_ref/")
visit_dates <- visit_dates %>%
  filter( visit_date != "" ) %>%
  select( record_id, redcap_event_name, visit_date )

### check counts: for how many patients we have all the follow-up dates (2 weeks, 6 weeks and 6 months?)
### in this case only for 12 of the patients all the information has been entered
counts <- visit_dates %>%
  group_by( record_id ) %>%
  mutate( dates_available = n( ) ) %>%
  select( record_id, dates_available ) %>%
  unique()
summary(as.factor( counts$dates_available))

### let's put together in a table all the information, combining the patient timeline with first admission and discharge date with the rest of follow-up dates when available
misc_followUp <- visit_dates %>%
  spread( redcap_event_name, visit_date ) %>%
  left_join( misc_mapping, by = c("record_id"="MUSIC_ID") ) %>%
  select( PATIENT_NUM = PAT_NUM, two_weeks = "2_weeks_postdischa_arm_1", 
          six_weeks = "6_weeks_postdischa_arm_1", six_months = "6_months_postdisch_arm_1")

head(misc_followUp)
```


Data transformation
- split `START_DATE` field into `date` and `time`
- perform a left join to combine extracted data with variables in lab dictionary
- reformat `ADMISSION_DATE` and `DISCHARGE_DATE` as dates
- calculate `days_after_admission` and `days_after_discharge`
And we join this information with the follow-up information. 
```{r}
subset_df$date <- sapply(strsplit( as.character(subset_df$START_DATE), " "), '[', 1)
subset_df$time <- sapply(strsplit( as.character(subset_df$START_DATE), " "), '[', 2)

finalTable <- left_join( subset_df, lab_dictionary, by = c("CONCEPT_CD" = "concept_cd") )

finalTable <- finalTable %>%
  mutate(date = dmy(date),
         ADMISSION_DATE = dmy(ADMISSION_DATE),
         DISCHARGE_DATE = dmy(DISCHARGE_DATE),
         days_after_admission = date - ADMISSION_DATE,
         days_after_discharge = date - DISCHARGE_DATE) %>%
  left_join( misc_followUp )
```

#### Check the quality of the follow up dates

```{r}
qc_dates <- finalTable %>%
  mutate( two_weeks_daysDiff = abs( ymd( two_weeks) - ymd( DISCHARGE_DATE)), 
          six_weeks_daysDiff = abs( ymd( six_weeks) - ymd( DISCHARGE_DATE)), 
          six_months_daysDiff = abs( ymd( six_months) - ymd( DISCHARGE_DATE)), 
          check_two_weeks = ifelse( two_weeks_daysDiff > 21 | two_weeks_daysDiff < 7, "check", "ok"), 
          check_six_weeks = ifelse( six_weeks_daysDiff > 49| six_weeks_daysDiff < 35, "check", "ok") , 
          check_six_months = ifelse( six_months_daysDiff > 214 | six_months_daysDiff < 150, "check", "ok")) %>%
  select( PATIENT_NUM, ADMISSION_DATE, DISCHARGE_DATE, two_weeks, six_weeks, 
          six_months, two_weeks_daysDiff, six_weeks_daysDiff, six_months_daysDiff, 
          check_two_weeks, check_six_weeks, check_six_months) %>%
  unique()
```



Extract `value` based on type of variable
- focus on `_date`, `_value`, `_unit`, and `_obtained` variables (identified as most prevalent in `data_dictionary_extraction.Rmd`)
  - choices accepted for `obtained` variables: 1, 0
  - choices accepted for `value` variables: no criteria. if NVAL_NUM is not NA, then NVAL_NUM. else TVAL_CHAR
  - choices accepted for `unit` variables: no criteria. use UNITS_CD column
  - choices accepted for `date` variables: use date column in mdy format
```{r}
# Create function to extract value
getLabValue <- function(row) {
  
  # for obtained variables, if the entry is present, the 'value' is automatically 1 except for:
  # APTT/PTT that choices are 2, APTT | 3, PTT | 0, Neither	
  if (row['vartype'] == 'obtained' ) {
    if( row['variableName'] == 'aptt_ptt_obtained'){
      value = '3' #in BCH PTT is the one estimated
    }else{
      value = '1'
    }
    
  # Variables with unit listed below the valaue field will be filled with ‘-88’
  # WBC and platelets 1, x10^3/uL | 2, 10^9/L | 3, Other --> for WBC and platelets is x10^3/uL			
  # triglycerid 1, mg/dL | 2, mmol/L | 3, Other	1	 --> for BCH the values mg/dL		
  # CRP 1, mg/L | 2, mg/dL | 3, Other --> for BCH the value is mg/dL
  # IgE: Expected unit value: 'mg/dL'; current unit value: 'unit/mL'
    
  } else if (row['vartype'] == 'unit') {
    if( row['variableName'] == 'crp_unit'){
      value = '2'
    }else if( row['variableName'] == 'triglycerid_unit'){
      value = '1'
    }else if( row['variableName'] == 'wbc_unit'){
      value = '1'
    }else if( row['variableName'] == 'platelet_unit'){
      value = '1'
    }else if (row['variableName'] == 'ige_unit'){
      value = 'unit/mL'
    } else {
      value = '-88'
    }
    
  # for date variables, use the date column
  # redcap requires the following format: 
  # (NOTE: Dates must be imported here only in M/D/Y format or Y-M-D format, regardless of the specific date format designated for this field.)

  } else if (row['vartype'] == 'date') {
    value = as.character(ymd(row['date']), format = "%m/%d/%Y")
    
  # for value variables, if NVAL_NUM is not NA, use NVAL_NUM. else use TVAL_CHAR
  # for igg and igm: return 1 Positive (detected) if >11, 3 Indeterminate if 9-11, 0 Not detected if <9
    
  } else if (row['vartype'] == 'value') {
    if(!is.na(row['NVAL_NUM'])){
      
      if(row['variableName'] %in% c('anti_sars_igg_value', 'anti_sars_igm_value')){
        if(as.numeric(row['NVAL_NUM']) > 11){
          value = '1'
        } else if(as.numeric(row['NVAL_NUM']) >= 9) {
          value = '3'
        } else {
          value = '0'
        } 
      } else {
        value = as.character(as.numeric(row['NVAL_NUM']))
      }
      
    } else {
      value = row['TVAL_CHAR']
    }
  }
    
  # for other variables, mark as 'UNKNOWN' and investigate; there should not be any of these for now
  else {
    value = 'UNKNOWN'
  }
   
  return(value)
}

```

```{r}
finalTable <- finalTable %>%
  # categorize the variable types
  mutate(vartype = gsub("^.*_", "", variableName)) %>%
  # remove entries with no data
  filter(!is.na(NVAL_NUM) | TVAL_CHAR != '') %>%
  # remove duplicated rows
  distinct %>%
  # filter out labs where reported value is "Not Reported" or "TNP"
  filter(!TVAL_CHAR %in% c("Not Reported", "TNP"))
         
finalTable$value <- apply(finalTable, 1, FUN = getLabValue)
```

lab_values_visit
1, Admission/first obtained during MIS-C hospitalization
2, Closest to discharge during MIS-C hospitalization
3, 2 Weeks Post Discharge
4, 6 Weeks Post Discharge
6, 6 Months Post Discharge


## Assign period based on the days of diff
```{r}
 finalTable <- finalTable %>%
   mutate(period = case_when(
          days_after_admission >= -2 & days_after_admission <=  2 ~ '1',
          days_after_discharge >= -3 & days_after_discharge <= 0 ~ '2', 
          days_after_discharge >= 7 & days_after_discharge < 21 ~ '3',
          days_after_discharge >= 21 & days_after_discharge <  63 ~ '4',
          days_after_discharge >= 63 & days_after_discharge < 365 ~ '6',
          TRUE ~ 'Out of period'))

#######
# Code to deal with multiple concept codes for lymphocytes and neutrophils
# We consolidate all concepts to one code
finalTable <- finalTable %>%
  mutate(CONCEPT_CD = gsub('LAB:3775664', 'LAB:3775715', CONCEPT_CD),
         CONCEPT_CD = gsub('LAB:3775663', 'LAB:3775714', CONCEPT_CD))

#######


# select which value to report when more than one lab value for the same patient and period 
finalTable <- finalTable %>%
  filter( period != "Out of period") %>%
  group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
  mutate( two_weeks_ref = if_else( !is.na( ymd(two_weeks) ), ymd(two_weeks), ymd( DISCHARGE_DATE ) + days(14) ), 
          six_weeks_ref = if_else( !is.na( ymd(six_weeks) ), ymd(six_weeks), ymd( DISCHARGE_DATE ) + days(42) ), 
          six_months_ref =if_else( !is.na( ymd(six_months)), ymd(six_months), ymd( DISCHARGE_DATE ) + days(182) ), 
          daysDiff = case_when(  period == 1 ~ as.character( abs(ymd(date) - ymd(ADMISSION_DATE))),
                                 period == 2 ~ as.character( abs(ymd(date) - ymd(DISCHARGE_DATE))),
                                 period == 3 ~ as.character( abs(ymd(date) - ymd(two_weeks_ref))),
                                 period == 4 ~ as.character( abs(ymd(date) - ymd(six_weeks_ref))),
                                  period == 6 ~ as.character( abs(ymd(date) - ymd(six_months_ref))),
                                TRUE ~ " "), 
          daysDiff = as.numeric( daysDiff ) )

finalTableFiltered <- finalTable %>%
 group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
  slice_min( order_by = daysDiff ) %>%
  slice_min( order_by = dmy_hms(START_DATE) )

```


Because we are choosing 1 lab value per time point (the closest to the date), we should not have more than one value for redcap_repeat_instance. 

```{r}
finalTableFiltered <- finalTableFiltered %>% 
  group_by(PATIENT_NUM) %>%
  mutate( lab_values_visit = period, 
    redcap_repeat_instance = lab_values_visit ) %>%
  arrange(redcap_repeat_instance)
```

Save the file as RData file
```{r}
latestDataUpdate <- finalTableFiltered
save(latestDataUpdate, file = "../local_ref/finalTableFilteredLatestUpdateData.RData")
```

Clean the environment
```{r}
rm(list=ls())
```


# Run the pipeline until the finalTableFiltered is created with the previous data extract and save it as an RData file
Import the lab data from MISC patients. Instead of importing a subset, we import the whole dataset (ag440_music_labs_nov21) and here we filter by the concepts that are in our lab_dictionary and we rename the admission and discharge date colum.  
```{r}
lab_dictionary <- read.delim("../../laboratory_values/data/labDictionary_updated.txt", header = TRUE, sep = ' ')

all_labs <- read.delim('../../laboratory_values/local_ref/ag440_music_labs_nov21.dsv')
subset_df <- all_labs %>%
  filter( CONCEPT_CD %in% lab_dictionary$concept_cd ) %>%
  rename('ADMISSION_DATE' = ADMISSIONDATE,
         'DISCHARGE_DATE' = DISCHARGE_DATE_FORMATTED)
```

Import the follow-up visit dates. The MUSIC_DATA_XXXX.csv is the file extracted from REDCap of the data previously entered by BCH experts. From there we will extract the specific follow-up dates for the MUSIC study for the MISC patients. The variable name that contains the date is "visit_date", extracted from the dictionary. The record id is the internal MUSIC identifier, and the redcap_event_name allows us to know which is the specific follow-up period the date is referring to. Because the ID are MUSIC Ids we also need to load the music_mapping_id. 

To create the music_id_mapping file we create this table in SQL:
create table music_id_mapping_feb22 as 
select m.MRN, m.MUSIC_ID, p.PAT_NUM from MUSIC_PATIENTLIST_FEB22 m, MRN_PATUUID_PATNUM p where m.MRN = p.MRN; 

```{r}
misc_mapping <- read.csv("..//local_ref/music_id_mapping_feb22.csv")
visit_dates  <- read.csv("../local_ref/MUSIC_DATA_2022-02-16_0900.csv")
visit_dates <- visit_dates %>%
  filter( visit_date != "" ) %>%
  select( record_id, redcap_event_name, visit_date )

### check counts: for how many patients we have all the follow-up dates (2 weeks, 6 weeks and 6 months?)
### in this case only for 12 of the patients all the information has been entered
counts <- visit_dates %>%
  group_by( record_id ) %>%
  mutate( dates_available = n( ) ) %>%
  select( record_id, dates_available ) %>%
  unique()
summary(as.factor( counts$dates_available))

### let's put together in a table all the information, combining the patient timeline with first admission and discharge date with the rest of follow-up dates when available
misc_followUp <- visit_dates %>%
  spread( redcap_event_name, visit_date ) %>%
  left_join( misc_mapping, by = c("record_id"="MUSIC_ID") ) %>%
  select( PATIENT_NUM = PAT_NUM, two_weeks = "2_weeks_postdischa_arm_1", 
          six_weeks = "6_weeks_postdischa_arm_1", six_months = "6_months_postdisch_arm_1")

head(misc_followUp)
```


Data transformation
- split `START_DATE` field into `date` and `time`
- perform a left join to combine extracted data with variables in lab dictionary
- reformat `ADMISSION_DATE` and `DISCHARGE_DATE` as dates
- calculate `days_after_admission` and `days_after_discharge`
And we join this information with the follow-up information. 
```{r}
subset_df$date <- sapply(strsplit( as.character(subset_df$START_DATE), " "), '[', 1)
subset_df$time <- sapply(strsplit( as.character(subset_df$START_DATE), " "), '[', 2)

finalTable <- left_join( subset_df, lab_dictionary, by = c("CONCEPT_CD" = "concept_cd") )

finalTable <- finalTable %>%
  mutate(date = dmy(date),
         ADMISSION_DATE = dmy(ADMISSION_DATE),
         DISCHARGE_DATE = dmy(DISCHARGE_DATE),
         days_after_admission = date - ADMISSION_DATE,
         days_after_discharge = date - DISCHARGE_DATE) %>%
  left_join( misc_followUp )
```

#### Check the quality of the follow up dates

```{r}
qc_dates <- finalTable %>%
  mutate( two_weeks_daysDiff = abs( ymd( two_weeks) - ymd( DISCHARGE_DATE)), 
          six_weeks_daysDiff = abs( ymd( six_weeks) - ymd( DISCHARGE_DATE)), 
          six_months_daysDiff = abs( ymd( six_months) - ymd( DISCHARGE_DATE)), 
          check_two_weeks = ifelse( two_weeks_daysDiff > 21 | two_weeks_daysDiff < 7, "check", "ok"), 
          check_six_weeks = ifelse( six_weeks_daysDiff > 49| six_weeks_daysDiff < 35, "check", "ok") , 
          check_six_months = ifelse( six_months_daysDiff > 214 | six_months_daysDiff < 150, "check", "ok")) %>%
  select( PATIENT_NUM, ADMISSION_DATE, DISCHARGE_DATE, two_weeks, six_weeks, 
          six_months, two_weeks_daysDiff, six_weeks_daysDiff, six_months_daysDiff, 
          check_two_weeks, check_six_weeks, check_six_months) %>%
  unique()
```



Extract `value` based on type of variable
- focus on `_date`, `_value`, `_unit`, and `_obtained` variables (identified as most prevalent in `data_dictionary_extraction.Rmd`)
  - choices accepted for `obtained` variables: 1, 0
  - choices accepted for `value` variables: no criteria. if NVAL_NUM is not NA, then NVAL_NUM. else TVAL_CHAR
  - choices accepted for `unit` variables: no criteria. use UNITS_CD column
  - choices accepted for `date` variables: use date column in mdy format
```{r}
# Create function to extract value
getLabValue <- function(row) {
  
  # for obtained variables, if the entry is present, the 'value' is automatically 1 except for:
  # APTT/PTT that choices are 2, APTT | 3, PTT | 0, Neither	
  if (row['vartype'] == 'obtained' ) {
    if( row['variableName'] == 'aptt_ptt_obtained'){
      value = '3' #in BCH PTT is the one estimated
    }else{
      value = '1'
    }
    
  # Variables with unit listed below the valaue field will be filled with ‘-88’
  # WBC and platelets 1, x10^3/uL | 2, 10^9/L | 3, Other --> for WBC and platelets is x10^3/uL			
  # triglycerid 1, mg/dL | 2, mmol/L | 3, Other	1	 --> for BCH the values mg/dL		
  # CRP 1, mg/L | 2, mg/dL | 3, Other --> for BCH the value is mg/dL
  # IgE: Expected unit value: 'mg/dL'; current unit value: 'unit/mL'
    
  } else if (row['vartype'] == 'unit') {
    if( row['variableName'] == 'crp_unit'){
      value = '2'
    }else if( row['variableName'] == 'triglycerid_unit'){
      value = '1'
    }else if( row['variableName'] == 'wbc_unit'){
      value = '1'
    }else if( row['variableName'] == 'platelet_unit'){
      value = '1'
    }else if (row['variableName'] == 'ige_unit'){
      value = 'unit/mL'
    } else {
      value = '-88'
    }
    
  # for date variables, use the date column
  # redcap requires the following format: 
  # (NOTE: Dates must be imported here only in M/D/Y format or Y-M-D format, regardless of the specific date format designated for this field.)

  } else if (row['vartype'] == 'date') {
    value = as.character(ymd(row['date']), format = "%m/%d/%Y")
    
  # for value variables, if NVAL_NUM is not NA, use NVAL_NUM. else use TVAL_CHAR
  # for igg and igm: return 1 Positive (detected) if >11, 3 Indeterminate if 9-11, 0 Not detected if <9
    
  } else if (row['vartype'] == 'value') {
    if(!is.na(row['NVAL_NUM'])){
      
      if(row['variableName'] %in% c('anti_sars_igg_value', 'anti_sars_igm_value')){
        if(as.numeric(row['NVAL_NUM']) > 11){
          value = '1'
        } else if(as.numeric(row['NVAL_NUM']) >= 9) {
          value = '3'
        } else {
          value = '0'
        } 
      } else {
        value = as.character(as.numeric(row['NVAL_NUM']))
      }
      
    } else {
      value = row['TVAL_CHAR']
    }
  }
    
  # for other variables, mark as 'UNKNOWN' and investigate; there should not be any of these for now
  else {
    value = 'UNKNOWN'
  }
   
  return(value)
}

```

```{r}
finalTable <- finalTable %>%
  # categorize the variable types
  mutate(vartype = gsub("^.*_", "", variableName)) %>%
  # remove entries with no data
  filter(!is.na(NVAL_NUM) | TVAL_CHAR != '') %>%
  # remove duplicated rows
  distinct %>%
  # filter out labs where reported value is "Not Reported" or "TNP"
  filter(!TVAL_CHAR %in% c("Not Reported", "TNP"))
         
finalTable$value <- apply(finalTable, 1, FUN = getLabValue)
```

lab_values_visit
1, Admission/first obtained during MIS-C hospitalization
2, Closest to discharge during MIS-C hospitalization
3, 2 Weeks Post Discharge
4, 6 Weeks Post Discharge
6, 6 Months Post Discharge


## Assign period based on the days of diff
```{r}
 finalTable <- finalTable %>%
   mutate(period = case_when(
          days_after_admission >= -2 & days_after_admission <=  2 ~ '1',
          days_after_discharge >= -3 & days_after_discharge <= 0 ~ '2', 
          days_after_discharge >= 7 & days_after_discharge < 21 ~ '3',
          days_after_discharge >= 21 & days_after_discharge <  63 ~ '4',
          days_after_discharge >= 63 & days_after_discharge < 365 ~ '6',
          TRUE ~ 'Out of period'))

#######
# Code to deal with multiple concept codes for lymphocytes and neutrophils
# We consolidate all concepts to one code
finalTable <- finalTable %>%
  mutate(CONCEPT_CD = gsub('LAB:3775664', 'LAB:3775715', CONCEPT_CD),
         CONCEPT_CD = gsub('LAB:3775663', 'LAB:3775714', CONCEPT_CD))

#######


# select which value to report when more than one lab value for the same patient and period 
finalTable <- finalTable %>%
  filter( period != "Out of period") %>%
  group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
  mutate( two_weeks_ref = if_else( !is.na( ymd(two_weeks) ), ymd(two_weeks), ymd( DISCHARGE_DATE ) + days(14) ), 
          six_weeks_ref = if_else( !is.na( ymd(six_weeks) ), ymd(six_weeks), ymd( DISCHARGE_DATE ) + days(42) ), 
          six_months_ref =if_else( !is.na( ymd(six_months)), ymd(six_months), ymd( DISCHARGE_DATE ) + days(182) ), 
          daysDiff = case_when(  period == 1 ~ as.character( abs(ymd(date) - ymd(ADMISSION_DATE))),
                                 period == 2 ~ as.character( abs(ymd(date) - ymd(DISCHARGE_DATE))),
                                 period == 3 ~ as.character( abs(ymd(date) - ymd(two_weeks_ref))),
                                 period == 4 ~ as.character( abs(ymd(date) - ymd(six_weeks_ref))),
                                  period == 6 ~ as.character( abs(ymd(date) - ymd(six_months_ref))),
                                TRUE ~ " "), 
          daysDiff = as.numeric( daysDiff ) )

finalTableFiltered <- finalTable %>%
 group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
  slice_min( order_by = daysDiff ) %>%
  slice_min( order_by = dmy_hms(START_DATE) )

```


Because we are choosing 1 lab value per time point (the closest to the date), we should not have more than one value for redcap_repeat_instance. 

```{r}
finalTableFiltered <- finalTableFiltered %>% 
  group_by(PATIENT_NUM) %>%
  mutate( lab_values_visit = period, 
    redcap_repeat_instance = lab_values_visit ) %>%
  arrange(redcap_repeat_instance)
```


Save the file as RData file
```{r}
previousDataUpdate <- finalTableFiltered
save(previousDataUpdate, file = "../local_ref/finalTableFilteredPreviousData.RData")
```

Clean the environment
```{r}
rm(list=ls())
```


## Load the two files saved for the data extracts to compare 

```{r}
### Simran approach

load("../local_ref/finalTableFilteredPreviousData.RData" )
load("../local_ref/finalTableFilteredLatestUpdateData.RData" )


p1 <- previousDataUpdate %>%
  select(PATIENT_NUM, variableName, period, value)

l1 <- latestDataUpdate %>%
  select(PATIENT_NUM, variableName, period, value)

combined <- full_join(p1, l1,
                      by = c('PATIENT_NUM', 'variableName', 'period'),
                      suffix = c('.prev', '.latest'))

na_prev <- combined %>% filter(is.na(value.prev))
na_latest <- combined %>% filter(is.na(value.latest))

combined2 <- combined %>%
  filter(!is.na(value.prev) & !is.na(value.latest)) %>%
  mutate(check = value.prev == value.latest)

# do the same count thing alba did before to account for multiple values
tocheck <- combined2 %>% filter(!check) %>%
  group_by(PATIENT_NUM, variableName, period) %>%
  mutate(count = n()) %>%
  filter( count < 2 )

# look at values only
tocheck_values <- tocheck %>%
  filter(grepl('value', variableName))

table(tocheck$PATIENT_NUM)
table(tocheck$period)

# the values which changed for patient 106495209 are the admission ones (MUSIC ID 120018) this one was admitted may 18 2020
# the values which changed for patient 114915319 are the 6 month ones (MUSIC ID 120051) this makes sense

patlist_may <- read.csv('../../local_ref/misc_patients_may.dsv')
patlist_feb <- read.csv('../../local_ref/music_patients_feb2022.csv')

```



```{r}
load("../local_ref/finalTableFilteredPreviousData.RData" )
load("../local_ref/finalTableFilteredLatestUpdateData.RData" )


latest <- latestDataUpdate %>%
  filter( vartype == "value") %>%
  select( PATIENT_NUM, CONCEPT_CD, period, latest_value = value) %>%
  unique()

previous <- previousDataUpdate %>%
  filter( vartype == "value") %>%
  select( PATIENT_NUM, CONCEPT_CD, period, previous_value = value) %>%
  unique()



overwrite <- full_join( previous, latest,  by=c("PATIENT_NUM", "CONCEPT_CD", "period"))

dataPreviouslyEntered <- overwrite %>%
  filter( ! is.na( previous_value ) )


dataPreviouslyEntered <- dataPreviouslyEntered %>%
  mutate( comparison = ifelse( previous_value == latest_value, "OK", "Overwrite"))

needToOverWrite <- dataPreviouslyEntered %>%
  filter( comparison == "Overwrite") %>%
  group_by( PATIENT_NUM, CONCEPT_CD, period ) %>%
  mutate(count = n()) %>%
  filter( count < 2 )


summary(as.factor( needToOverWrite$period )) # 32 in period 1 and 6
length(unique( needToOverWrite$PATIENT_NUM )) # 2 patients

#percentage
round( nrow( needToOverWrite )/nrow( dataPreviouslyEntered )*100, 2)


nasCases <- dataPreviouslyEntered %>%
  filter( is.na(comparison))

length(unique(nasCases$PATIENT_NUM))


### trying to understand why we have multiple values for the same period... need to work more on it and review the previous estimation
latest2 <- latestDataUpdate %>%
  filter( vartype == "value") %>%
  select( PATIENT_NUM, CONCEPT_CD, period, START_DATE, latest_value = value) %>%
  unique()

previous2 <- previousDataUpdate %>%
  filter( vartype == "value") %>%
  select( PATIENT_NUM, CONCEPT_CD, period, START_DATE, previous_value = value) %>%
  unique()


overwrite2 <- full_join( previous2, latest2,  by=c("PATIENT_NUM", "CONCEPT_CD", "period", "START_DATE"))

dataPreviouslyEntered2 <- overwrite2 %>%
  filter( ! is.na( previous_value ) )


dataPreviouslyEntered2 <- dataPreviouslyEntered2 %>%
  mutate( comparison = ifelse( previous_value == latest_value, "OK", "Overwrite"))

needToOverWrite2 <- dataPreviouslyEntered2 %>%
  filter( comparison == "Overwrite") %>%
  group_by( PATIENT_NUM, CONCEPT_CD, period, START_DATE ) %>%
  mutate(count = n()) %>%
  filter( count < 2 )

summary(as.factor( needToOverWrite2$period )) # 32 in period 1 and 6
length(unique( needToOverWrite2$PATIENT_NUM )) # 2 patients

#percentage
round( nrow( needToOverWrite2 )/nrow( dataPreviouslyEntered2 )*100, 2)


nasCases2 <- dataPreviouslyEntered2 %>%
  filter( is.na(comparison2))

length(unique(nasCases2$PATIENT_NUM))

```

