---
title: 'Data refresh: internal QC'
author: "Simran Makwana & Alba Gutierrez"
date: "02/24/2022"
output: html_document
---


# Aim
To compare the latest data extract with the previous extraction to make sure the data is consistent. 

# Library Installation
```{r}
library(tidyverse)
library(lubridate)
library(generics)
```

# Labs data: QC


Import the lab data from MISC patients. Instead of importing a subset, we import the whole dataset (ag440_music_labs_nov21) and here we filter by the concepts that are in our lab_dictionary and we rename the admission and discharge date colum.  
```{r}
##all data
misc_mapping <- read.csv("..//local_ref/music_id_mapping_feb22.csv")

all_labs_latest_update <- read.delim('../local_ref/ag440_music_labs_feb22.dsv') %>%
  mutate( DATE = sapply(strsplit( as.character(START_DATE), " "), '[', 1), 
          combData = paste( PATIENT_NUM, CONCEPT_CD, START_DATE, TVAL_CHAR, NVAL_NUM, UNITS_CD, sep = "**")) %>%
  unique()

all_labs_previous_extract <- read.delim('../../laboratory_values/local_ref/ag440_music_labs_nov21.dsv') %>%
    filter( PATIENT_NUM %in% all_labs_latest_update$PATIENT_NUM ) %>%
  mutate( DATE = sapply(strsplit( as.character(START_DATE), " "), '[', 1), 
          combData = paste( PATIENT_NUM, CONCEPT_CD, START_DATE, TVAL_CHAR, NVAL_NUM, UNITS_CD, sep = "**"))%>%
  filter( dmy(DATE) <= "2021-05-30") %>%
  unique()
  
all_labs_latest_update <- all_labs_latest_update %>%
  filter( dmy(DATE) <= "2021-05-30" & 
            PATIENT_NUM %in% all_labs_previous_extract$PATIENT_NUM )

# look for same exact rows
dim( all_labs_previous_extract )
dim( all_labs_latest_update)

#if any distinct data use this 
xx <- all_labs_latest_update[ ! all_labs_latest_update$combData %in% all_labs_previous_extract$combData, ]
yy <- all_labs_previous_extract[ ! all_labs_previous_extract$combData %in% all_labs_latest_update$combData, ]

nrow(xx)
nrow(yy)
```


