---
title: 'ETL: Medications During Hospitalization eCRF'
author: "Simran Makwana"
date: "9/16/2021"
output: html_document
---

# This notebook contains the following steps:
1.Load data extract for the concepts of interest
2. Determine expected values for REDCap 
3. Calculate medcur1_current column
4. Transform and format

# 0. Libraries
```{r}
library(tidyverse)
library(lubridate)
```

Here we specify for which patients we want to run the pipeline for.
```{r}
music_patients_admission_after_nov21 <- read.csv("../../laboratory_values/local_ref/texasChildren/misc_med_data_observations.csv")

#redcap <- read.csv("../local_ref/redcap_output_medications_during.csv") %>%
#  pull( record_id ) %>%
#  unique()

### we must additionally make sure we are not running the pipeline on any patients who were not discharged before the data refresh
#REFRESH_DATE_CUTOFF <- mdy('11-01-2021')

#newPatients <- music_patients_admission_after_nov21 %>%
#  filter( ! MUSIC.ID %in% redcap,
#          mdy(discharge_date) < REFRESH_DATE_CUTOFF)

newPatients <- music_patients_admission_after_nov21$patient_num #run pipeline on all patients in tch - no data refresh


# now we have our list of new poatient IDs we want to run the pipeline on.

```


# 1. Load data extract for the codes of interest
Use the same query as medications_before_and_after -- see that pipeline for details
```{r}
# we flter the subset df to only include the patients of interest
#misc_mapping <- read.csv("../../dataRefresh/local_ref/MUSIC_ID_MAPPING") %>%
#  rename(PATIENT_NUM = PAT_NUM) %>%
#  filter( MUSIC_ID %in% newPatients$MUSIC.ID )

#df <- read.delim('../../dataRefresh/local_ref/ag440_music_meds_feb22.dsv') %>%
#  filter(PATIENT_NUM %in% misc_mapping$PATIENT_NUM) #####

df <- music_patients_admission_after_nov21

# if the code is formatted like so: '2.00', convert to '02.00'
fix_code_formatting <- function(x) {
  ifelse(test = is.na(x), yes = NA, 
         no = ifelse(test = !grepl("[[:digit:]]{2}.", x), yes = (paste0('0', x)), no = (x)))
}

datadict <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv') %>%
  filter(Form.Name == 'additional_medications_during_hospitalization') %>%
  select(Variable...Field.Name, Choices..Calculations..OR.Slider.Labels) %>%
  separate(Variable...Field.Name, c('a', 'b', 'c'), sep = '_') %>%
  filter(!b %in% c('na', 'type', 'name', 'start', 'ae')) %>%
  mutate(a = str_replace(a, '[[:digit:]]+', ''),
         Variable...Field.Name = paste0(a, 'X', '_', b, '_', c)) %>%
  unique() %>% 
  separate_rows(Choices..Calculations..OR.Slider.Labels, sep = '\\|') %>%
  mutate(ECRF_CODE = str_replace(Choices..Calculations..OR.Slider.Labels, ',.*$', ''),
         ECRF_CODE = str_trim(ECRF_CODE),
         ECRF_CODE = fix_code_formatting(ECRF_CODE)) %>%
  select(-Choices..Calculations..OR.Slider.Labels) %>% 
  mutate(ECRF_CODE = fix_code_formatting(ECRF_CODE))

#bch_med_concept_summary <- read_csv('../../medications_before_and_after/data/bch_med_concept_summary_toReview.csv') %>%
#  filter(RXNORM_DESCS != 'greater than 5 RXNORM codes mapped to BCH code' | (CONCEPT_CD == 'ADMINMED:2797919' & ECRF_CODE == '07.08'))

# why did we remove this in BCH? CONCEPT_CD == 'ADMINMED:2797919' & ECRF_CODE == '07.08' heparin
tch_med_concept_summary <- read.delim('../../medications_before_and_after/data/tch_to_ecrf_med_mapping.txt')

#bch_med_concept_summary <- left_join(bch_med_concept_summary, datadict, by = 'ECRF_CODE')

###TCH SPECIFIC: ADD ADDITIONAL CODES
unassigned_meds <- df %>% select(concept_cd, med_route) %>% unique()
write.csv(unassigned_meds, '../tch_unassigned_meds.csv', quote = FALSE)
# ecrf has been searched for remaining codes and updated in tch_unassigned_meds_withcodes.csv
unassigned_meds_withcodes <- read_csv('../tch_unassigned_meds_withcodes.csv') %>%
  select(-med_route) %>% 
  filter(!CONCEPT_CD %in% tch_med_concept_summary$concept_cd) %>%
  unique() %>%
  filter(!is.na(CONCEPT_CD))


tch_med_concept_summary <- tch_med_concept_summary %>% rbind(data.frame('keyword' = 'manual', 
                                                                        'ECRF_CODE' = unassigned_meds_withcodes$ECRF_CODE, 
                                                                        'concept_cd' = unassigned_meds_withcodes$CONCEPT_CD)) 


#df1 <- df %>%
#  left_join(unassigned_meds_withcodes %>% select(ECRF_CODE, CONCEPT_CD), by = 'CONCEPT_CD') %>%
#  mutate(ECRF_CODE = ifelse(ECRF_CODE.x == "99.99" & !is.na(ECRF_CODE.y), ECRF_CODE.y, ECRF_CODE.x)) %>%
#  select(-ECRF_CODE.x, -ECRF_CODE.y, -X, -X.1)


#df <- left_join(df, bch_med_concept_summary %>% 
#                  select(ECRF_CODE, CONCEPT_CD, Variable...Field.Name), by = 'CONCEPT_CD') 
df <- left_join(df %>% 
                  rename(CONCEPT_CD=concept_cd), 
                tch_med_concept_summary %>% 
                  mutate(ECRF_CODE = fix_code_formatting(ECRF_CODE)) %>% 
                  select(ECRF_CODE, CONCEPT_CD=concept_cd), by = 'CONCEPT_CD') %>%
  left_join(datadict  %>% select(ECRF_CODE, Variable...Field.Name))

# NAs that appear are not present in our mapping / are not requested by the ECRF
# we assign them the 99.99 code for 'other'
df <- df %>%
  mutate(ECRF_CODE = ifelse(is.na( ECRF_CODE ),
                            '99.99',
                            ECRF_CODE))

# assumptions: 
# - INOUT_CD = INPATIENT
# - med_specification = GIVEN
#df <- df %>% 
#  filter(inout_cd == 'INPATIENT',
#         med_specification == 'Given')

```


```{r}

# we have some concepts which we want to remove, even from the 'other' mappings
#concepts_toremove <- read.csv('../../medications_before_and_after/data/bch_concepts_to_remove.csv') %>% pull(CONCEPT_CD)


# for 'other' medications, load in bch concept_dimension table for medications.
# we onlyh report the names for unknown medications.
# even though we are reporting all 'other' medications, we will remove the general sodium chloride concepts as they are not informative.
#bch_concept_dim <- read.delim('../../medications_before_and_after/data/concept_dimensionMeds.dsv') %>%
#  filter(!CONCEPT_CD %in% concepts_toremove) %>%
#  select(CONCEPT_CD, NAME_CHAR) %>%
#  # we have to manually add the concept for 'albumin human', based on clinician review with Simran Mahanta and Audery Dionne
#  add_row(CONCEPT_CD = 'ADMINMED:2797321', NAME_CHAR = 'albumin human') %>%
#  unique %>%
#  group_by(CONCEPT_CD) %>% 
#  summarise(NAME_CHAR = paste0(NAME_CHAR, collapse = ', '))



df <- df %>% 
  #left_join(bch_concept_dim, by = 'CONCEPT_CD') %>%
  mutate(medhospX_name = ifelse(ECRF_CODE == '99.99',
                                tolower(CONCEPT_CD),
                                '')) %>%
  filter(!is.na(medhospX_name))


```


```{r}
# which medications do we report dates for?
report_dates <- read.csv('../../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv') %>%
  filter(Form.Name == 'additional_medications_during_hospitalization', 
         Variable...Field.Name == 'medhosp1_start_dt') %>%
  pull(Branching.Logic..Show.field.only.if....)

report_dates

# we write the following regular expression to capture all of these codes:
report_dates <- '26.|28.|99.99|09.99|09.03|09.09|18.03|18.04|18.06|18.09|18.10'
```

# 2. Filter to medication values during hospitalization

```{r}
misc_patient <- read.csv('../local_ref/../../laboratory_values/local_ref/texasChildren/misc_patient.csv') %>%
  rename(ADMISSION_DATE = Admission.Date,
         DISCHARGE_DATE = Discharge.Date,
         patient_num = MUSIC.Patient.ID) %>%
  mutate(patient_num = as.character(patient_num))

df <- left_join(df, misc_patient)

df <- df %>%
  rename(START_DATE=start_date,) %>%
  mutate(#date = dmy(sapply(strsplit( as.character(START_DATE), " "), '[', 1)),
#         DISCHARGE_DATE = dmy(sapply(strsplit( as.character(END_DATE), " "), '[', 1)),
         #DISCHARGE_DATE = dmy(sapply(strsplit( as.character(DISCHARGE_DATE), " "), '[', 1)),
         #ADMISSION_DATE = dmy(sapply(strsplit( as.character(ADMISSIONDATE), " "), '[', 1)),
    date = ymd(date), 
    DISCHARGE_DATE = as.Date( DISCHARGE_DATE, format = "%m/%d/%y"), 
    ADMISSION_DATE = as.Date( ADMISSION_DATE, format = "%m/%d/%y"),
         days_after_admission = as.numeric(date - ADMISSION_DATE),
         days_after_discharge = as.numeric(date - DISCHARGE_DATE)) %>%
  filter(days_after_admission >= 0,
         days_after_discharge <= 0)# %>%
  # here we also remove lidocaine concepts that are not IV.
  #filter(!(CONCEPT_CD == 'ADMINMED:4103284' & TVAL_CHAR != 'ID'))

```

# Add additional columns
- medhospX_type: overall category of medication
- medhospX_start_dt: start date, mm-dd-yyyy *****This does not appear to be needed
- medhospX_name: name of medication, only include for 'other' medications
- med_num refers to the 'X' in the variable name. (internal)
- redcap_repeat_instance refers to the form number. you can have 10 med_nums per form.
```{r}

# only report the dates for the specified medications. for all other medications, medhosp_start_dt should be blank
# for medications where we are reporting the date, only report the 'start date', or the earliest date.
df2 <- df %>% 
  rename(PATIENT_NUM = patient_num)  %>%
  mutate(medhospX_start_dt = format(date, "%Y-%m-%d")) %>%
  group_by(PATIENT_NUM, ECRF_CODE, medhospX_name) %>%
  summarise(medhospX_start_dt = ifelse(grepl(report_dates, ECRF_CODE),
                                    min(medhospX_start_dt),
                                    ''),
            CONCEPT_CD = first(CONCEPT_CD),
            Variable...Field.Name = first(Variable...Field.Name)) %>%
  ungroup() %>%
  unique()



df3 <- df2 %>%
  group_by(PATIENT_NUM) %>%
  arrange(ECRF_CODE) %>%
  mutate(rownum = row_number(),
         med_num = ifelse(rownum %% 10 == 0, 10, rownum %% 10),
         redcap_repeat_instance = (rownum - 1) %/% 10 + 1,
         medhospX_type = ifelse(ECRF_CODE == '99.99',
                                '99.99',
                                (str_replace(str_replace(ECRF_CODE, '\\...', ''), '^0', '')))) %>%
  ungroup() %>%
  select(-rownum) 

# rename X in variable name depending on 'med_num'

df4 <- df3 %>% 
  mutate(Variable...Field.Name = str_replace(Variable...Field.Name, 'X', as.character(med_num)))



```

# Pivot into final format
```{r}
# pivot variable names
out <- pivot_wider(df4,
                  id_cols = c(PATIENT_NUM, medhospX_type, medhospX_name, redcap_repeat_instance, med_num, medhospX_start_dt),   
                  names_from = Variable...Field.Name,
                  values_from = ECRF_CODE,
                  values_fill = NA) 

# pivot medhospX_start_dt
out2 <- out %>%
  mutate(medhospX_start_dt_names = str_replace('medhospX_start_dt', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_start_dt_names,
    values_from = medhospX_start_dt,
    values_fill = NA)
  

# pivot medhospX_type
out3 <- out2 %>%
  mutate(medhospX_type_names = str_replace('medhospX_type', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_type_names,
    values_from = medhospX_type,
    values_fill = NA)

# pivot medhospX_name
out4 <- out3 %>%
  mutate(medhospX_name_names = str_replace('medhospX_name', 'X', as.character(med_num))) %>%
  pivot_wider(
    names_from = medhospX_name_names,
    values_from = medhospX_name,
    values_fill = NA) %>%
  select(-med_num)


# map BCH ids to MUSIC IDs for REDCAP
# id_mapping <- read.csv('../../local_ref/music_id_mapping.csv')
#id_mapping <- misc_mapping
out5 <- out4 %>% #left_join(out4, id_mapping %>% select(PATIENT_NUM, MUSIC_ID)) %>% 
  ungroup() %>% 
  rename(record_id = PATIENT_NUM) %>%
  #rename(record_id = MUSIC_ID) %>%
  mutate(redcap_event_name = "day_1_arm_1",
         redcap_repeat_instrument = "additional_medications_during_hospitalization") %>%
  relocate(record_id, redcap_event_name, redcap_repeat_instrument) 
  
# collapse rows by form
out6 <- out5 %>%
  group_by(record_id, redcap_repeat_instance) %>%
  summarise_all(~first(na.omit(.)))
#out5 <- out4 %>%
#  group_by(record_id, redcap_repeat_instance) %>%
#  summarise_each(funs(first(.[!is.na(.)])))

# remove NA column introduced by 'other' variables
out7 <- out6 %>% select(-grep('NA', colnames(out6)))


```


```{r}
meds_data_point <- out5 %>%
  select(-redcap_repeat_instance, -redcap_event_name, -redcap_repeat_instrument)
meds_data_point_filtered <- meds_data_point %>%
  tidyr::pivot_longer( cols = c(2:ncol(meds_data_point)), 
                       names_to = "variable", 
                       values_to = "value") %>%
  dplyr::filter( value != "")

```


```{r}
### write test output for test patients 120002 and 120003
#write.csv(out7 %>%
#            filter(record_id %in% c(120006, 120002, 120095, 120123)) %>%
#            mutate(record_id = ifelse(record_id == 120006,
#                                      888006, 
#                                      ifelse(record_id == 120002,
#                                             888002,
#                                             ifelse(record_id == 120095,
#                                                    888095,
#                                                    888123)))),
#          '../local_ref/redcap_TEST_output_medications_during.csv',
#          row.names = FALSE, 
#          na = '')

# write output
#write.csv(out7, '../local_ref/redcap_output_medications_during.csv', row.names = FALSE, na = '')
write.csv(out7, '../local_ref/tch_medications_during_092023.csv', row.names = FALSE, na = '')
#write.csv(out7 %>% filter(record_id %in% c(120122, 120123, 120124)),
#          '../local_ref/TEST_DATA_REFRESH_NEW_PATIENTS_MAR22_redcap_output_medications_during.csv',
#          row.names = FALSE, na = '')

```




