---
title: "Generate Potential eCRF to RXNORM Mapping"
author: "Alba Gutierrez"
date: "6/15/2021"
output: html_document
---

# Load Packages
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
```

# Load UMLS database subset (all RXNORM codes)
The following SQL was used to generate the subset:
create table ag440_rxnorm as
select CUI,SAB, TTY, CODE, STR from mrconso where sab = 'RXNORM'and lat = 'ENG';

```{r}
rxnorm_subset <- read.delim("../data/rxnorm.dsv")
```

# Load the medication table specifications
The full list with all the medications including code, generic name, class and trades, is obtained from the MUSIC dropbox. The original file is called "CodeListD(Meds) MUSIC 9-7-20". It is a word document that has been transformed into a csv format. 
```{r}
meds_keywords <- read.csv("../data/CodeListD(Meds) MUSIC 9-7-20.csv", colClasses = "character" )
print(length(unique(meds_keywords$Code)))
```

There are 32 medication classes, divided into 349 codes for 325 generic medication names. For each code, for some we have the specific trades (from 1 to 5 distinct trades), and for others we only have the generic name and class. 

We will focus on those medications that we can extract in an automatic way, based on the information we have. We will use as keyword for each medication code, the generic names that are distinct than "Not specified" and distinct than "Other...", and we will use as keywords the generic names and the trades. 
```{r}
meds_auto <- meds_keywords %>%
  dplyr::filter( tolower(Generic.Name) != "not specified" & 
                 ! grepl( "^other",tolower(Generic.Name)))

print(length(unique(meds_auto$Code)))
```
We have 295 medications that we can query in an automatic way. 

# From eCRF to RXNORM
Search through the UMLS subset for the keywords associated with the eCRF terms.
We use as keywords the Trades and the Generic Name for our search.
We are not restricting by TTY in this eCRF.

```{r}
# transform empty Trades into NAs
#meds_auto <- meds_auto %>%
#  mutate(across(everything(), ~ifelse(.=="", NA, as.character(.))))

# add keywords to meds_auto df
meds_auto <- meds_auto %>% mutate(key = str_c(Generic.Name, Trade.1, Trade.2, Trade.3, Trade.4, Trade.5, sep = '|'),  # this line combines the generic name with the trade names, separated by |, to form a single keyword
                                  key = str_remove(key, '\\|+$'), # this line removes any trailing pipes from the keyword
                                  key = tolower(key)) # this line converts all keywords to lowercase

# create a blank dataframe for ecrf_to_rxnorm
ecrf_to_rxnorm <- data.frame('ECRF_CODE' = '', 'ECRF_GENERIC_NAME' = '', 'RXNORM_CODE' = '', 'RXNORM_DESCRIPTION' = '')
ecrf_to_rxnorm <- ecrf_to_rxnorm[-1,]

count = 0
for( i  in 1:nrow( meds_auto ) ){
  #print( i )
  #keyword <- tolower(paste0( "*", meds_auto[i,] %>%
  #  unite( "keywords", Generic.Name, Trade.1:Trade.5, 
  #         na.rm = TRUE, sep = ".*|") %>%
  #  select( keywords), "*"))
  
  rxnorm_output <- rxnorm_subset %>% 
    dplyr::filter( grepl( keyword, tolower(STR))) %>%       # search for keyword within rxnorm STR (lowercase)
    dplyr::group_by( CODE ) %>%     # group by code, as some codes are listed multiple times with different TTY or STR
    dplyr::summarise(RXNORM_DESCRIPTION = paste(STR, collapse = ' | ')) %>%   # combining all STRs to retain all information 
    dplyr::mutate( ECRF_CODE = code, 
                   ECRF_GENERIC_NAME = meds_auto %>% filter(Code == code) %>% pull(Generic.Name), 
                   RXNORM_CODE = as.character( CODE )) %>%
    dplyr::select( ECRF_CODE, ECRF_GENERIC_NAME, RXNORM_CODE, RXNORM_DESCRIPTION)

  # if the keyword matches 1 or more RXNORM entries, add the results to ecrf_to_rxnorm
  if( nrow(rxnorm_output ) > 0){
      ecrf_to_rxnorm <- rbind( ecrf_to_rxnorm, rxnorm_output )
  }
  
  # else, add 1 to the count of non-mapping keywords, and print the keyword for reference
  else{ 
    count = count + 1
    print( keyword )
  }
  
}
```

```{r}
print(count)
print(length(unique(ecrf_to_rxnorm$ECRF_CODE)))
print(length(unique(ecrf_to_rxnorm$RXNORM_CODE)))
```

There are only 19 codes for which we cannot find any RXNORM that matches in UMLS.
We have 276 (93.56%) of medication codes found that mapped to 53,722 RXNORM codes.

# From RXNOMR to BCH
Check which of the RXNORM are present in the BCH internal mapping. 
In the BCH mapping we have 2,196 RXNORM codes that map to 19,141 BCH internal codes.
```{r}
bchmapping <- read.csv("MedicationsMapBCH.csv")
length(unique(bchmapping$ACT_CONCEPT_CD))
length(unique(bchmapping$BCH_CONCEPT_CD))

bchmapping$RXNORM_CODE <- sapply(strsplit( as.character(bchmapping$ACT_CONCEPT_CD), "RXNORM:"), '[', 2)

rxnorm_to_bch <- inner_join( ecrf_to_rxnorm, bchmapping, by="RXNORM_CODE")
length( unique( rxnorm_to_bch$ECRF_CODE))
length( unique( rxnorm_to_bch$RXNORM_CODE))
length( unique( rxnorm_to_bch$BCH_CONCEPT_CD))
```

In the BCH mapping we find 199 eCRF medication codes that map to 632 RXNORM codes and to 7,538 BCH internal codes.


# From BCH mapping to MISC patients
We summarise the information of the codes that are present in the MISC patients. 
```{r}
misc_meds <- read.delim("./music_patients_meds.dsv")

misc_meds <- misc_meds %>% 
  dplyr::filter( CONCEPT_CD %in%  rxnorm_to_bch$BCH_CONCEPT_CD )

misc_meds_info <- misc_meds %>% 
  dplyr::inner_join( rxnorm_to_bch, by=c("CONCEPT_CD"="BCH_CONCEPT_CD"))

length(unique(misc_meds_info$ECRF_CODE))
length(unique(misc_meds_info$CONCEPT_CD))
```
We find 120 eCRF medication codes present in the BCH MISC patients. 
We make a summary to see how many MISC patients present each one, and in which date ranges. 

First we add a column based on when the medication was prescribed, during the hospitalization, or before or after. 
```{r}
misc_meds_info$med_date <- sapply(strsplit( as.character(misc_meds_info$START_DATE), "[ ]"), '[', 1)
misc_meds_info$misc_start_date <- sapply(strsplit( as.character(misc_meds_info$ADMISSIONDATE), "[ ]"), '[', 1)
misc_meds_info$misc_end_date <- sapply(strsplit( as.character(misc_meds_info$END_DATE), "[ ]"), '[', 1)
misc_meds_info$when <- ifelse( dmy( misc_meds_info$med_date ) < dmy( misc_meds_info$misc_start_date), "before", ifelse( dmy( misc_meds_info$med_date) > dmy( misc_meds_info$misc_end_date), "after", "during"))
```

As a sanity check, and to better understand the data, we will create two subsets, one with the "during" hospitalization, and one with the "before or after", and we check if for the "during hospitalization all the medications prescribed are "ADMINMED". 

```{r}
misc_meds_info$prefix <- sapply(strsplit( as.character(misc_meds_info$CONCEPT_CD), "[:]"), '[', 1)
during <- misc_meds_info %>%
  dplyr::filter( when == "during")

summary( as.factor( during$prefix))

before_after <- misc_meds_info %>%
  dplyr::filter( when != "during")
summary( as.factor( before_after$prefix))
```

And we create the summary for each medication in our dataset. 

```{r}
bch_med_concept_summary <- misc_meds_info %>%
  dplyr::group_by(CONCEPT_CD) %>%
  dplyr::summarise( min_date = min( dmy( med_date ) ), 
                    max_date = max( dmy( med_date ) ), 
                    distinct_patients = length(unique(PATIENT_NUM)), 
                    distinct_observations = length(PATIENT_NUM), 
                    form_name = "medications", 
                    ecrf_code = ECRF_CODE, 
                    ecrf_generic_name = ECRF_GENERIC_NAME, 
                    bch_concept_path = BCH_CONCEPT_PATH, 
                    bch_concept_description = BCH_NAME_CHAR) %>%
  dplyr::select( form_name, ecrf_code, ecrf_generic_name, CONCEPT_CD,
                 bch_concept_path, bch_concept_description, min_date, max_date, 
                 distinct_patients, distinct_observations)

bch_med_concept_summary <- unique( bch_med_concept_summary )
```

```{r}
ecrf_med_concept_summary <- misc_meds_info %>%
  dplyr::group_by( ECRF_CODE ) %>%
  dplyr::summarise( min_date = min( dmy( med_date ) ), 
                    max_date = max( dmy( med_date ) ), 
                    distinct_patients = length(unique(PATIENT_NUM)), 
                    distinct_observations = length(PATIENT_NUM),
                    distinct_concepts = length(unique(CONCEPT_CD)),
                    form_name = "medications", 
                    ecrf_generic_name = ECRF_GENERIC_NAME) %>%
  dplyr::select( form_name, ECRF_CODE, ecrf_generic_name,
                 min_date, max_date, 
                 distinct_patients, distinct_observations, distinct_concepts)

ecrf_med_concept_summary <- unique( ecrf_med_concept_summary )

```

# Save the output summary

```{r}
write.csv(bch_med_concept_summary, './bch_med_concept_summary_toReview.csv', row.names = FALSE)
write.csv(ecrf_med_concept_summary, './ecrf_med_concept_summary_toReview.csv', row.names = FALSE)
```

