---
title: "Concordance estimation: manual entry vs. automatic entry"
author: "Alba Gutiérrez-Sacristán & Simran Makwana"
date: "2023-09-22"
output: html_document
---

# Load libraries
```{r}
rm(list=)
library(dplyr)
library(tidyr)
library(eulerr)
```

# Laboratory values
## Load the files
We need 3 files as input:
- The TCH automatic file generated with our etl pipeline
- The TCH manual file, downloaded from REDCap with the manual data entered
- The MUSIC data dictionary
```{r}
auto_tch <- read.csv('../laboratory_values/local_ref/lab_output_texas_092023.csv', colClasses ="character")
#auto_tch <- read.csv('../laboratory_values/local_ref/lab_output_texas_092023_testInpatient_other.csv', colClasses ="character")

manual_tch <- read.csv("../laboratory_values/local_ref/texasChildren/MUSIC_DATA_2023-09-01_1508.csv", colClasses =  "character") %>%
  filter( record_id %in% auto_tch$record_id ) #filtering by the common patients

# MUSIC data dictionary from the general common_ref folder
datadict <- read.csv('../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv')
```

## Laboratory values selection
```{r}
lab_datadict <- datadict %>% 
  dplyr::filter(Form.Name == "laboratory_values") 

auto_tch <- auto_tch %>%
  dplyr::filter( redcap_repeat_instrument == "laboratory_values")

manual_tch <- manual_tch %>%
  dplyr::filter( redcap_repeat_instrument == "laboratory_values")
```

## Identify the data in-scope
We define the data in scope as those variables that can be automatically extractable, and the we can then compare in the manual vs. automatic data entry. 

### Reformat the data
Create an id based on the patient ID and the visit value and pivot the data to have tables with 3 columns:
- id 
- variable
- value

```{r}
totalManualVar <- manual_tch %>%
  dplyr::select( c("record_id", lab_datadict$Variable...Field.Name) ) 

totalManualVar <- totalManualVar %>%
  tidyr::pivot_longer( cols = c(2:ncol( totalManualVar)), 
                       names_to = "variable", 
                       values_to = "value_manual") %>%
  dplyr::filter( ! value_manual %in% c("", " ")) 
nrow( totalManualVar)

manual_tch <- manual_tch %>%
  dplyr::select( c("record_id", lab_datadict$Variable...Field.Name) ) %>%
  dplyr::mutate( id=paste0( record_id, "-",lab_values_visit)) %>%
  dplyr::select( -record_id, - lab_values_na, - lab_values_visit ) 

manual_tch <- manual_tch %>% # remove columns that are not needed
  tidyr::pivot_longer( cols = c(1:ncol( manual_tch) - 1), 
                       names_to = "variable", 
                       values_to = "value_manual") %>%
  dplyr::filter( ! value_manual %in% c("", " ")) 

auto_tch <-  auto_tch %>%
  dplyr::mutate( id=paste0( record_id, "-",lab_values_visit)) %>%
  dplyr::select( -record_id, -lab_values_visit, - redcap_event_name, - redcap_repeat_instrument,-redcap_repeat_instance ) # remove columns that are not needed

auto_tch <- auto_tch %>%
  tidyr::pivot_longer( cols = c(1:ncol(auto_tch)-1), 
                       names_to = "variable", 
                       values_to = "value_auto") %>%
  dplyr::filter( value_auto != "") 

```
Identify the variables that are in scope (auto extract possible). 
```{r}
nrow(manual_tch)
manual_tch_in_scope <- manual_tch %>%
  dplyr::filter( variable %in% auto_tch$variable )
nrow(manual_tch_in_scope)

manual_tch_Not_in_scope <- manual_tch %>%
  dplyr::filter( ! variable %in% auto_tch$variable )

toCheck <- as.data.frame( table( manual_tch_Not_in_scope$variable))
```
## Get the concordance 
We merge the manual_tch_in_scope with the auto_tch one, and we estimate for how many cases the value is the same in both. 
```{r}
manualVsauto <- left_join( manual_tch_in_scope, auto_tch, by =c("id", "variable"))
```

We #re-format dates changing NA auto values to 0 when manual is 0 and auto is NA and removing .0 from 4.0, 13.0 etc. 
We also remove white spaces and make sure everything is comparable. 
Finally we do the comparison to assign which cases are the same and which ones are different.
```{r}
#manualVsauto2 <- manualVsauto
manualVsauto <- manualVsauto %>%
  dplyr::mutate( value_manual = as.character( trimws( value_manual)), 
                 value_manual = ifelse( grepl( "\\.0$", value_manual) == TRUE, 
                                        gsub("\\.0", "", value_manual), value_manual ),
                 value_auto = as.character( trimws( value_auto)), 
                 value_auto = gsub("<", "", value_auto ),
                 value_auto = ifelse( grepl( "\\.0$", value_auto) == TRUE, 
                                        gsub("\\.0", "", value_auto), value_auto ))

manualVsauto <- manualVsauto %>%
  dplyr::mutate( type = sapply(strsplit( variable, "_"), tail, 1),
                 #value_auto =  #ifelse( type == "date",  
                               #as.character( as.Date( value_auto, format = "%m/%d/%Y")),
                               #value_auto), 
                 value_manual = as.character( trimws( value_manual)), 
                 value_auto = as.character( trimws( value_auto)), 
                 value_auto = ifelse( value_manual == 0 & is.na(value_auto), 0, value_auto ),
                 concordance = ifelse( value_manual == value_auto, "same", "different"))
```

Get a summary of the results
```{r}
results <- summary(as.factor( manualVsauto$concordance))
results
round(as.numeric(100* results["same"]/nrow(manualVsauto)), 2)

different <- manualVsauto %>% filter( concordance == "different")
missing <- manualVsauto %>% filter( is.na(concordance) )
different <- different %>% mutate( patient = sapply(strsplit( id, "-"), '[', 1), visit = sapply(strsplit( id, "-"), '[', 2))
obtained <- different %>% filter( type == "obtained" & value_auto == 0)
```
Plot the overlap
```{r}
totalA = sum(results)
totalB = nrow(auto_tch)
common = as.numeric(results["same"])
vd <- euler(c(manual=totalA, auto=totalB, "manual&auto"= common))
plot(vd, legend = TRUE, quantities =  list(type = c("counts", "percent")), fills = c("lightblue2", "lightyellow"))
```


# Medication values 

## Load the files
We need 3 files as input:
- The TCH automatic file generated with our etl pipeline
- The TCH manual file, downloaded from REDCap with the manual data entered
- The MUSIC data dictionary
```{r}
rm(list=ls())
auto_tch <- read.csv('../medications_during/local_ref/tch_medications_during_092023.csv', colClasses ="character")

# tch manual file downloaded from REDCap
manual_tch <- read.csv("../laboratory_values/local_ref/texasChildren/MUSIC_DATA_2023-09-01_1508.csv", colClasses =  "character") %>%
  filter( record_id %in% auto_tch$record_id ) #filtering by the common patients


# MUSIC data dictionary from the general common_ref folder
datadict <- read.csv('../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv')
```

## Medication values selection
```{r}
datadict <- datadict %>% 
  dplyr::filter(Form.Name == "additional_medications_during_hospitalization") 

manual_tch <- manual_tch %>%
  dplyr::filter( redcap_repeat_instrument == "additional_medications_during_hospitalization")

auto_tch <- auto_tch %>%
  dplyr::filter( redcap_repeat_instrument == "additional_medications_during_hospitalization")
```

## Transform into table that are comparables
```{r}
auto_tch <-  auto_tch %>%
  dplyr::select( - redcap_event_name, - redcap_repeat_instrument ) # remove columns that are not needed

auto_tch <- auto_tch %>%
  tidyr::pivot_longer( cols = c(3:ncol(auto_tch)), 
                       names_to = "variable", 
                       values_to = "value_auto") %>%
  dplyr::filter( value_auto != "") 
print(paste0("Total medication values extracted automatically: ", nrow( auto_tch)))


auto_tch_toCompare <- auto_tch %>%
  dplyr::mutate( mednum = sapply(strsplit( variable, "_"), head, 1),
                 medtype =  sapply(strsplit( variable, "_"), '[', 2))%>%
  dplyr::group_by( record_id, redcap_repeat_instance, mednum ) %>%
  dplyr::arrange(factor(medtype, levels = c("type", "code", "name", "start"))) %>%
  dplyr::summarise( counts = n(), 
                    new_val = ifelse( counts == 1, value_auto, paste(value_auto, collapse="*")))

manual_tch <- manual_tch %>%
  dplyr::select( c("record_id", "redcap_repeat_instance",  datadict$Variable...Field.Name) ) %>%
  dplyr::select( - med_na )

manual_tch <- manual_tch %>% # remove columns that are not needed
  tidyr::pivot_longer( cols = c(3:ncol(manual_tch)), 
                       names_to = "variable", 
                       values_to = "value_auto") %>%
  dplyr::filter( value_auto != "" & value_auto != 0 ) 

print(paste0("Total medication values entered manually: ", nrow( manual_tch)))


manual_tch_toCompare <- manual_tch %>%
  dplyr::mutate( mednum = sapply(strsplit( variable, "_"), head, 1),
                 medtype =  sapply(strsplit( variable, "_"), '[', 2))%>%
  dplyr::group_by( record_id, redcap_repeat_instance, mednum ) %>%
  dplyr::arrange(factor(medtype, levels = c("type", "code", "name", "start"))) %>%
  dplyr::summarise( counts = n(), 
                    new_val = ifelse( counts == 1, value_auto, paste(value_auto, collapse="*")))

```

## Merge auto and manual extraction to get the performance

```{r}
# merge both files by the id we have created and the variable name
manual <- manual_tch_toCompare %>%
  dplyr::ungroup() %>%
  dplyr::mutate( id = paste0( record_id, "###", new_val )) %>%
  dplyr::select( record_id, manual_val = new_val, manual_counts= counts, id ) %>%
  unique()

auto <- auto_tch_toCompare %>%
  dplyr::ungroup() %>%
  dplyr::mutate( id = paste0( record_id, "###", new_val )) %>%
  dplyr::select( record_id, auto_val = new_val, auto_counts = counts, id ) %>%
  unique()

manualVsauto <- full_join( manual, auto, by =c("id", "record_id")) %>%
  dplyr::mutate( auto_counts = ifelse( is.na(auto_counts), 0, auto_counts ), 
                 manual_counts = ifelse( is.na(manual_counts), 0, manual_counts ), 
                total_counts = ifelse( manual_counts == auto_counts, auto_counts, 
                                        ifelse( manual_counts == 0, auto_counts, manual_counts))) %>%
  dplyr::select( record_id, id, manual_val, auto_val, total_counts )

manualNotAuto <- manual[ ! manual$id %in% auto$id, ]
manualAndAuto <- manual[ manual$id %in% auto$id, ]

print(paste0( "Total medication values entered manually:", sum(manual$manual_counts))) 
print(paste0( "Concordance: ", sum(manualAndAuto$manual_counts))) 
```

## Investigate the .99 meds
```{r}
otherMeds <- manualNotAuto %>%
  dplyr::mutate( others = ifelse( grepl("\\.99", manual_val) == TRUE, "other" , "non")) %>%
  dplyr::filter(others == "other" ) %>%
  dplyr::select(manual_val, manual_counts ) %>%
  unique()

medsVals <- otherMeds %>%
  dplyr::mutate( med_description = ifelse( manual_counts == 4, 
                                           sapply(strsplit( manual_val, "[*]"), '[', 3), 
                                           ifelse( manual_counts == 3 & grepl("^99.99",manual_val)== TRUE, sapply(strsplit( manual_val, "[*]"), '[', 2),
                                                   sapply(strsplit( manual_val, "[*]"), '[', 3))) ) %>% dplyr::select( med_description ) %>% unique()

## save the values and look for the equivalent that we have in the EHR
write.table( medsVals, file = "../medications_during/medsFreeText.txt", col.names = TRUE, row.names = FALSE, quote = FALSE, sep = "\t")
```

We mapped the meds that we found that we codified as 99.99 to the same name that they entered manually. We also found a typo on their end for Doxycyline, that is Doxycycline. We will update that on the manual description of the variables for further comparison. 
Not all the meds they entered as .99 we have in the 99.99 category, some as Remdesivir or Enalapril, had their own category different than .99. 


# Session Information Summary
```{r}
sessionInfo()
```
