---
title: "Concordance estimation: manual entry vs. automatic entry"
author: "Alba Gutiérrez-Sacristán & Simran Makwana"
date: "2023-09-22"
output: html_document
---

# Load libraries
```{r}
rm(list=)
library(dplyr)
library(tidyr)
```

# Load the files
We need 3 files as input:
- The TCH automatic file generated with our etl pipeline
- The TCH manual file, downloaded from REDCap with the manual data entered
- The MUSIC data dictionary
```{r}
auto_tch <- read.csv('../laboratory_values/local_ref/lab_output_texas_092023.csv', colClasses ="character")
#auto_tch <- read.csv('../laboratory_values/local_ref/lab_output_texas_092023_testInpatient_other.csv', colClasses ="character")

manual_tch <- read.csv("../laboratory_values/local_ref/texasChildren/MUSIC_DATA_2023-09-01_1508.csv", colClasses =  "character") %>%
  filter( record_id %in% auto_tch$record_id ) #filtering by the common patients

# MUSIC data dictionary from the general common_ref folder
datadict <- read.csv('../common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv')
```

# Laboratory values selection
```{r}
lab_datadict <- datadict %>% 
  dplyr::filter(Form.Name == "laboratory_values") 

auto_tch <- auto_tch %>%
  dplyr::filter( redcap_repeat_instrument == "laboratory_values")

manual_tch <- manual_tch %>%
  dplyr::filter( redcap_repeat_instrument == "laboratory_values") %>%
  dplyr::select(record_id, lab_datadict$Variable...Field.Name)
```

# Identify the data in-scope
We define the data in scope as those variables that can be automatically extractable, and the we can then compare in the manual vs. automatic data entry. 

## Reformat the data
Create an id based on the patient ID and the visit value and pivot the data to have tables with 3 columns:
- id 
- variable
- value

```{r}
manual_tch <- manual_tch %>%
  dplyr::select( c("record_id", lab_datadict$Variable...Field.Name) ) %>%
  dplyr::mutate( id=paste0( record_id, "-",lab_values_visit)) %>%
  dplyr::select( -record_id, - lab_values_na, - lab_values_visit ) 

manual_tch <- manual_tch %>% # remove columns that are not needed
  tidyr::pivot_longer( cols = c(1:ncol( manual_tch) - 1), 
                       names_to = "variable", 
                       values_to = "value_manual") %>%
  dplyr::filter( ! value_manual %in% c("", " ")) 

auto_tch <-  auto_tch %>%
  dplyr::mutate( id=paste0( record_id, "-",lab_values_visit)) %>%
  dplyr::select( -record_id, -lab_values_visit, - redcap_event_name, - redcap_repeat_instrument,-redcap_repeat_instance ) # remove columns that are not needed

auto_tch <- auto_tch %>%
  tidyr::pivot_longer( cols = c(1:ncol(auto_tch)-1), 
                       names_to = "variable", 
                       values_to = "value_auto") %>%
  dplyr::filter( value_auto != "") 

```

Identify the variables that are in scope (auto extract possible). 
```{r}
nrow(manual_tch)
manual_tch_in_scope <- manual_tch %>%
  dplyr::filter( variable %in% auto_tch$variable )
nrow(manual_tch_in_scope)

manual_tch_Not_in_scope <- manual_tch %>%
  dplyr::filter( ! variable %in% auto_tch$variable )

toCheck <- as.data.frame( table( manual_tch_Not_in_scope$variable))
```
# Get the concordance 
We merge the manual_tch_in_scope with the auto_tch one, and we estimate for how many cases the value is the same in both. 
```{r}
manualVsauto <- left_join( manual_tch_in_scope, auto_tch, by =c("id", "variable"))
```

We #re-format dates changing NA auto values to 0 when manual is 0 and auto is NA and removing .0 from 4.0, 13.0 etc. 
We also remove white spaces and make sure everything is comparable. 
Finally we do the comparison to assign which cases are the same and which ones are different.
```{r}
#manualVsauto2 <- manualVsauto
manualVsauto <- manualVsauto %>%
  dplyr::mutate( value_manual = as.character( trimws( value_manual)), 
                 value_manual = ifelse( grepl( "\\.0$", value_manual) == TRUE, 
                                        gsub("\\.0", "", value_manual), value_manual ),
                 value_auto = as.character( trimws( value_auto)), 
                 value_auto = gsub("<", "", value_auto ),
                 value_auto = ifelse( grepl( "\\.0$", value_auto) == TRUE, 
                                        gsub("\\.0", "", value_auto), value_auto ))

manualVsauto <- manualVsauto %>%
  dplyr::mutate( type = sapply(strsplit( variable, "_"), tail, 1),
                 #value_auto =  #ifelse( type == "date",  
                               #as.character( as.Date( value_auto, format = "%m/%d/%Y")),
                               #value_auto), 
                 value_manual = as.character( trimws( value_manual)), 
                 value_auto = as.character( trimws( value_auto)), 
                 value_auto = ifelse( value_manual == 0 & is.na(value_auto), 0, value_auto ),
                 concordance = ifelse( value_manual == value_auto, "same", "different"))
```

Get a summary of the results
```{r}
results <- summary(as.factor( manualVsauto$concordance))
results
round(as.numeric(100* results["same"]/nrow(manualVsauto)), 2)

different <- manualVsauto %>% filter( concordance == "different")
missing <- manualVsauto %>% filter( is.na(concordance) )
different <- different %>% mutate( patient = sapply(strsplit( id, "-"), '[', 1), visit = sapply(strsplit( id, "-"), '[', 2))
obtained <- different %>% filter( type == "obtained" & value_auto == 0)
```


# Session Information Summary
```{r}
sessionInfo()
```
