---
title: "Generate Summary Table for eCRF Variables"
author: "Simran Makwana"
date: "6/30/2021"
output: html_document
---

# Packages and files
```{r}
library(tidyverse)

datadict <- read.csv('common_ref/MUSIC_DataDictionary_V3_4Dec20_final version_clean_0.csv')

datadict_labs <- datadict %>%
  filter(Form.Name == 'laboratory_values')
datadict_meds <- datadict %>%
  filter(Form.Name == 'medications_before_and_after_hosp')

loinc_mapping <- read.csv('laboratory_values/data/eCRF_to_LOINC_forReview.csv')

meds_codeslist <- read.csv('medications_before_and_after/data/CodeListD(Meds) MUSIC 9-7-20.csv')
med_mapping <- read.csv('medications_before_and_after/data/')
```

# Summarize number of variables in data dictionary for laboratory_values and medications_before and after

## Lab values summary

### Total number of variables
Distinct laboratory variable defined as a distinct *prefix*
Aoart from this, of the rows in the data dictionary, there are several which will not be reported in the final output CSV and therefore should be excluded.
- lab_values_na
- *_date
- anything where Filed Type == 'calc'
```{r}
datadict_labs <- datadict_labs %>%
  filter(Variable...Field.Name != 'lab_values_na',
         !grepl('.*_date', Variable...Field.Name),
         Field.Type != 'calc', 
         !grepl('.*other.*', Variable...Field.Name)) %>%
  mutate(prefix = gsub(Variable...Field.Name, pattern = '_obtained|_date|_value|_man|_other|_unit', replacement = '')) 

datadict_labs %>% pull(prefix) %>% n_distinct()
  

```

### Number of variables available for auto extraction
We define variables available for auto extraction as those with identifiable eCRF to *LOINC* mappings. 
```{r}
loinc_mapping %>% pull(eCRF.TERM.DESCRIPTION) %>% n_distinct()

```

### Number of variables available for auto extraction at BCH
We define BCH specific auto extraction variables as those with identifiable eCRF to LOINC to BCH concept mappings.
```{r}
# 64 - based on MUSIC Laboratory Values Form: mapping from eCRF numeric variables and BCH variables

```

### Number of variables which we can extract information automatically for MIS-C patients
We define these variables as variables which we are able to extract data from at least one MIS-C patient from BCH. 
```{r}
# 61 - based on MUSIC Laboratory Values Form: mapping from eCRF numeric variables and BCH variables

```

## Medications summary 

### Total number of variables
Derived directly from meds_codeslist
```{r}
meds_codeslist %>% nrow()

```
### Number of variables available for auto extraction
We define variables available for auto extraction as those with **potentially** identifiable eCRF to *RXNORM* mappings. 
Note that we might not have found an RXNORM code for all of these medications yet but that they could be expected to be automatically extracted to some extent. 
Remove the following as they require further manual review:
- Code ending in .99 (other)
- Code ending in .00 (not specified)

*This number does not reflect those which map to RXNORM yet*
Refer to eCRF_to_RXNORM_mapping_potential.Rmd
```{r}
meds_codeslist <- meds_codeslist %>%
  filter(!grepl('.99$|.00$', Code))

meds_codeslist %>% nrow()

```









